---
title: "StabMap: Generate Breast Cancer CITE-seq dataset object"
author: "Shila Ghazanfar"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
       html_document:
                     toc: true
                     toc_float:
                           collapsed: false
                           smooth_scroll: false
                     code_folding: hide
                     fig_width: 10 
                     fig_height: 8
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE,
                      cache = FALSE, cache.lazy = FALSE)
```

```{r}
set.seed(2021)
```

Load scripts and packages.

```{r}
source("../scripts/initialise.R")
library(igraph)
library(gtools)
library(DropletUtils)
library(CiteFuse)
```

Data is from Wu et al (https://www.nature.com/articles/s41588-021-00911-1),
where CITE-seq data from one patient is extracted, CID3838, corresponding
to HER2 case.

RNA modality was downloaded from GEO via supplementary file https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSM5354514
and saved locally into BreastCancerCITE/RNA folder. File name was renamed from
count_matrix_barcodes.tsv to barcodes.tsv; from count_matrix_genes.tsv to genes.tsv;
and from count_matrix_sparse.mtx to matrix.mtx.

ADT modality was downloaded from Broad Institute portal
https://singlecell.broadinstitute.org/single_cell/study/SCP1039
and save locally into BreastCancerCITE/ADT folder.

# Load RNA data

```{r}
rna = read10xCounts(samples = "../../BreastCancerCITE/RNA/CID3838")

rna_meta = read.csv("../../BreastCancerCITE/RNA/CID3838/metadata.csv", header = TRUE,
                    row.names = 1)[rna$Barcode,]

colData(rna) <- DataFrame(rna_meta)

rna
```

# Load ADT data

Note that there are different numbers of cells identified, so we collect the 
cells detected in both modalities.

```{r}
adt = read10xCounts(samples = "../../BreastCancerCITE/ADT/CID3838/umi_count/")
adt$Barcode <- paste0("CID3838_", adt$Barcode)
colnames(adt) <- adt$Barcode

cite_prots = toupper(gsub("-...............$", "", rownames(adt)))
cite_prots <- setNames(cite_prots, cite_prots)

mapping = rbind(
  c("CD31.VWF", "CD31"),
  c("CD8", "CD8A"),
  c("HER2..3B5.", "HER2"),
  c("PD.1", "PD-1")
)
colnames(mapping) = c("IMC", "CITE")

cite_prots_new = cite_prots
cite_prots_new[mapping[,"CITE"]] <- mapping[,"IMC"]
cite_prots_new <- as.character(cite_prots_new)

rownames(adt) <- cite_prots_new

table(colnames(adt) %in% colnames(rna))
table(colnames(rna) %in% colnames(adt))

adt
```

# Generate a joint object using CiteFuse

```{r}
cells = intersect(colnames(rna), colnames(adt))
length(cells)

cite_list = list(
  "RNA" = counts(rna[,cells]),
  "ADT" = counts(adt[,cells])
)

sce = preprocessing(cite_list)
colData(sce) <- colData(rna)[cells,]

sce
```

```{r}
saveRDS(sce, file = "../../BreastCancerCITE/cite_sce.Rds")
```

# Finish

```{r}
sessionInfo()
```