---
title: "StabMap: Generate Breast Cancer IMC dataset"
author: "Shila Ghazanfar"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
       html_document:
                     toc: true
                     toc_float:
                           collapsed: false
                           smooth_scroll: false
                     code_folding: hide
                     fig_width: 10 
                     fig_height: 8
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE,
                      cache = FALSE, cache.lazy = FALSE)
```

```{r}
set.seed(2021)
```

Load scripts and packages.

```{r}
source("../scripts/initialise.R")
library(igraph)
library(gtools)
library(fst)
```

Data is downloaded from zenodo link https://zenodo.org/record/6036188#.Y2Cu8exBxqs 
from Danenberg et al (https://www.nature.com/articles/s41588-022-01041-y),
saved locally into BreastCancerIMC folder

# Load clinical data

We subset patients according to ER+ and HER2 status.

```{r}
clin = read_fst("../../BreastCancerIMC/IMCClinical.fst")

mb_keep = subset(clin, ERStatus == "pos" & 
                   PAM50 == "HER2" & 
                   !is.na(PAM50))$metabric_id

length(mb_keep)
```

# Load IMC data

Subset the IMC detected cells to those that correspond to tumours and belong
to the selected patients.

```{r}
imc_raw = read.csv("../../BreastCancerIMC/SingleCells.csv", header = TRUE, row.names = NULL)

imc = subset(imc_raw,
             metabric_id %in% mb_keep & 
               is_tumour == 1)

rownames(imc) <- paste0(imc$metabric_id, "_", imc$ImageNumber, "_", imc$ObjectNumber)

dim(imc)
```

# Create SCE object for samples

Split proteins from other variables

```{r}
not_prots = c("ImageNumber","ObjectNumber","metabric_id", "cellPhenotype",
              "is_epithelial", "is_tumour" ,"is_normal", "is_dcis",
              "is_interface","is_perivascular", "is_hotAggregate",
              "DNA1","DNA2","Location_Center_X", "Location_Center_Y",
              "AreaShape_Area") 

imc_prots = setdiff(colnames(imc), not_prots)

imc_exprs = imc[,imc_prots]
colnames(imc_exprs) <- toupper(imc_prots)

sce = SingleCellExperiment(assays = list(logcounts = t(imc_exprs)),
                           colData = imc[,not_prots],
                           reducedDims = list(coords = imc[,c("Location_Center_X", "Location_Center_Y")]))

sce
```

```{r}
saveRDS(sce, file = "../../BreastCancerIMC/imc_sce.Rds")
```

# Finish

```{r}
sessionInfo()
```