---
title: "StabMap simulation: Mouse Gastrulation Data"
author: "Shila Ghazanfar"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
       html_document:
                     toc: true
                     toc_float:
                           collapsed: false
                           smooth_scroll: false
                     code_folding: hide
                     fig_width: 10 
                     fig_height: 8
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE,
                      cache = FALSE, cache.lazy = FALSE)
```

```{r}
set.seed(2021)
```

Load scripts and packages.

```{r}
library(MouseGastrulationData)
library(scattermore)
library(ggpubr)
source("../scripts/initialise.R")
```

Helper plotting function:

```{r}
pl = function(embedding_UMAP, name, coltype = "type", facet = FALSE) {
  ind = sample(rownames(embedding_UMAP))
  df = data.frame(cell = ind,
                  U1 = embedding_UMAP[ind,1],
                  U2 = embedding_UMAP[ind,2],
                  type = factor(type[ind]),
                  ctype = ct_all[ind])
  df$col = df[,coltype]
  p = ggplot(df, aes(x = U1, y = U2, colour = col)) + 
    # geom_point(size = 0.3) + # replaced with scattermore
    geom_scattermore(pointsize = 2.3) +
    theme_classic() + 
    theme(legend.position = "none") +
    theme(axis.text = element_blank()) +
    theme(axis.ticks = element_blank()) +
    xlab("") +
    ylab("") +
    ggtitle(name)
  
  if (facet) {
    p <- p + facet_wrap(~type, nrow = 2, scales = "free")  +
      NULL
  }
  
  if (coltype == "ctype") {
    p <- p + scale_colour_manual(values = celltype_colours)
  }
  
  if (coltype == "type") {
    p <- p + scale_colour_manual(values = c("Query" = "red", "Reference" = "black"))
  }
  
  return(p)
}
```

Load data via the MouseGastrulationData Bioconductor package.

```{r}
mt = MouseGastrulationData::AtlasSampleMetadata
samples = mt[mt[, "stage"] == "E8.5", "sample"]

atlas = EmbryoAtlasData(type = "processed", samples = samples)
atlas <- logNormCounts(atlas)

decomp <- modelGeneVar(atlas)
hvgs <- rownames(decomp)[decomp$mean > 0.05 & decomp$total > 0]
length(hvgs)
atlas <- atlas[hvgs,]

celltype_colours = MouseGastrulationData::EmbryoCelltypeColours

ct_all = atlas$celltype
names(ct_all) <- colnames(atlas)

table(atlas$sample)
```

Build simulation set up. Treat three samples as the reference data, and 
treat the fourth sample as a query dataset with fewer features - selected
randomly among the entire set of features. Perform StabMap embedding, and 
compare against PCA, UINMF, and MultiMAP.

Assess performance by predicting cell types of the query dataset, using the 
reference dataset cells as the training set, and reporting the classification 
accuracy of the query dataset.

```{r}
nGenes_all = c(100, 200, 500, 1000, nrow(atlas),
               100, 200, 500, 1000,
               100, 200)

names(nGenes_all) <- paste0("Sim_", seq_len(length(nGenes_all)))

labels = "celltype"
assayNameReference = "logcounts"
assayNameQuery = "logcounts"
doUMAP = TRUE
nPCs = 50

resFile = "../../output/MGD_StabMap_example/MGD_StabMap_example_res.Rds"
simGenesFile = "../../output/MGD_StabMap_example/MGD_StabMap_example_simGenes.Rds"

if (!file.exists(resFile)) {
  res = NULL
} else {
  res = readRDS(resFile)
}

if (!file.exists(simGenesFile)) {
  simGenes = list()
} else {
  simGenes = readRDS(simGenesFile)
}

for (sim in names(nGenes_all)) {
  
  nGenes = nGenes_all[sim]
  if (is.null(simGenes[[sim]])) {
    genes = sample(rownames(atlas)[rowVars(assay(atlas, assayNameReference)) > 0])[seq_len(nGenes)]
    simGenes[[sim]] <- genes
  } else {
    genes <- simGenes[[sim]]
  }
  
  for (querySample in samples) {
    
    for (referenceSample in setdiff(samples,querySample)) {
      
      print(nGenes)
      print(paste0("Query: ", querySample))
      print(paste0("Reference: ", referenceSample))
      
      # check has this already been done?
      if (any(res[,"Sim"] == sim & 
              res[,"querySample"] == querySample & 
              res[,"referenceSample"] == referenceSample)) next
      
      referenceSCE = atlas[,atlas$sample %in% referenceSample]
      
      querySCE = atlas[genes, atlas$sample %in% querySample]
      
      # data type factor
      type = factor(ifelse(c(colnames(referenceSCE), colnames(querySCE)) %in% colnames(referenceSCE),
                           "Reference", "Query"))
      names(type) <- c(colnames(referenceSCE), colnames(querySCE))
      
      # input files as lists
      SCE_list = list(Reference = referenceSCE, Query = querySCE)
      assayNames = list(Reference = assayNameReference, Query = assayNameQuery)
      assay_list = mapply(assay, SCE_list, assayNames)
      counts_list = mapply(assay, SCE_list, "counts")
      labels_list = list(Reference = setNames(referenceSCE$celltype, colnames(referenceSCE)))
      
      g = mosaicDataUpSet(assay_list)
      print(g)
      pdf(file = paste0("../../Figures/raw/MGD_", 
                        sim, "_", 
                        querySample, "_",
                        referenceSample,
                        "_upset.pdf"), height = 3, width = 5)
      print(g)
      dev.off()
      
      # perform PCA
      PC_embedding = mapPCA(assay_list,nPCs = nPCs)
      dim(PC_embedding)
      
      # perform UINMF
      UINMF_embedding = UINMF_wrap(counts_list = counts_list)
      dim(UINMF_embedding)
      
      StabMap_embedding = stabMap(assay_list,
                                  reference_list = "Reference",
                                  ncomponentsReference = nPCs,
                                  ncomponentsSubset = nPCs,
                                  projectAll = TRUE)
      dim(StabMap_embedding)
      
      MultiMAP_embedding = MultiMAP_wrap(assay_list, verbose = TRUE)
      dim(MultiMAP_embedding)
      
      embeddings_names = c(
        "PCA" = "PC_embedding",
        "UINMF" = "UINMF_embedding",
        "MultiMAP" = "MultiMAP_embedding",
        "StabMap" = "StabMap_embedding"
      )
      
      nobatch = c("MultiMAP_embedding", "UINMF_embedding")
      noumap = c("MultiMAP_embedding")
      
      cnames = unlist(lapply(assay_list, colnames))
      batchFactor_ref <- ifelse(cnames %in% colnames(referenceSCE),"Reference", "Query")
      batchFactor = as.character(interaction(batchFactor_ref, colData(atlas)[cnames, "sample"]))
      names(batchFactor) <- cnames
      
      sapply(embeddings_names, function(nm) {
        print(nm)
        # can include optional flags for certain embeddings to 
        # not be passed through a batch correction stage
        # e.g. from MultiMAP
        if (nm %in% nobatch) {
          # i.e. do nothing
          print(paste0("no batch correction for ", nm))
          assign(paste0(nm, "_corrected"), get(nm), envir = .GlobalEnv)
        } else {
          assign(paste0(nm, "_corrected"), reducedMNN_batchFactor(get(nm), batchFactor),
                 envir = .GlobalEnv)
        }
      }, simplify = FALSE)
      
      embeddings_corrected_UMAP = sapply(embeddings_names, function(nm) {
        if (nm %in% noumap) return(get(nm)) # i.e. do nothing
        calculateUMAP_rnames(get(paste0(nm, "_corrected")))
      }, simplify = FALSE)
      
      p_corrected = 
        wrap_plots(
          c(
            mapply(pl, embeddings_corrected_UMAP, names(embeddings_corrected_UMAP), SIMPLIFY = FALSE),
            mapply(pl, embeddings_corrected_UMAP, names(embeddings_corrected_UMAP), MoreArgs = list(coltype = "ctype"), SIMPLIFY = FALSE)
          )) +
        plot_layout(nrow = 2, ncol = length(embeddings_corrected_UMAP), byrow = TRUE)
      print(p_corrected)
      
      
      # Plot the UMAPs
      gList = mapply(pl, embeddings_corrected_UMAP, names(embeddings_corrected_UMAP), MoreArgs = list(coltype = "ctype", facet = TRUE), SIMPLIFY = FALSE)
      
      g = wrap_plots(gList) + plot_layout(nrow = 1)
      print(g)
      
      g = wrap_plots(gList[c("StabMap", "PCA", "MultiMAP", "UINMF")]) + plot_layout(nrow = 1)
      print(g)
      
      gList2 = mapply(pl, embeddings_corrected_UMAP, names(embeddings_corrected_UMAP),
                      MoreArgs = list(coltype = "type", facet = FALSE), SIMPLIFY = FALSE)
      g = wrap_plots(gList2) + plot_layout(nrow = 1)
      print(g)
      
      g = wrap_plots(c(lapply(gList[c("StabMap", "PCA", "MultiMAP", "UINMF")],
                              function(x) x + 
                                theme(strip.background = element_rect(colour=NA, fill=NA)) + 
                                theme(plot.title = element_text(hjust = 0.5))
      ),
      lapply(gList2[c("StabMap", "PCA", "MultiMAP", "UINMF")],
             function(x) x + ggtitle(NULL)))) + 
        plot_layout(nrow = 2, heights = c(2,1))
      
      print(g)
      ggsave(g, file = paste0("../../Figures/raw/MGD_", sim, "_", querySample,
                              "_", referenceSample,
                              "_umaps.pdf"), height = 6, width = 9)
      
      g_leg = as_ggplot(get_legend(gList[[1]] + theme(legend.position = "right")
                                   + guides(colour = guide_legend(title = "",
                                                                  ncol = 2,
                                                                  override.aes = list(size = 6.5)))))
      ggsave(g_leg, file = paste0("../../Figures/raw/MGD_celltype_umaps_leg.pdf"), height = 5, width = 4)
      
      g_leg_type = as_ggplot(get_legend(gList2[[1]] + theme(legend.position = "right")
                                        + guides(colour = guide_legend(title = "",
                                                                       ncol = 3,
                                                                       override.aes = list(size = 6)))))
      ggsave(g_leg_type, file = paste0("../../Figures/raw/MGD_type_umaps_leg.pdf"), height = 1, width = 4)
      
      # save the embeddings as Rdata objects
      save(PC_embedding, StabMap_embedding, PC_embedding_corrected,
           StabMap_embedding_corrected, UINMF_embedding, MultiMAP_embedding, 
           embeddings_corrected_UMAP,
           file = paste0("../../output/MGD_StabMap_example/MGD_StabMap_example_embeddings_",
                         sim, "_", querySample, "_", referenceSample,
                         ".RData"))
      
      # predict cell type labels using knn with k = 5
      referenceLabels = colData(referenceSCE)[,labels]
      names(referenceLabels) = colnames(referenceSCE)
      
      queryLabels = colData(querySCE)[,labels]
      names(queryLabels) = colnames(querySCE)
      
      # calculate accuracy of the query cells
      embeddings_accuracy = sapply(embeddings_names, function(nm) {
        print(nm)
        
        # only use cells with labels
        data_all = get(paste0(nm, "_corrected"))
        labels_train = referenceLabels[!is.na(referenceLabels) & names(referenceLabels) %in% rownames(data_all)]
        
        knn_out = embeddingKNN(data_all,
                               labels_train,
                               type = "uniform_fixed",
                               k_values = 5)
        
        acc = mean(isEqual(knn_out[names(queryLabels),"predicted_labels"], queryLabels), na.rm = TRUE)
        
        return(acc)
      }, simplify = FALSE)
      
      print(embeddings_accuracy)
      
      # calculate the resubstitution accuracy
      embeddings_resub_accuracy = sapply(embeddings_names, function(nm) {
        print(nm)
        
        # only use cells with labels
        data_all = get(paste0(nm, "_corrected"))
        labels_train = referenceLabels[!is.na(referenceLabels) & names(referenceLabels) %in% rownames(data_all)]
        
        knn_out = embeddingKNN(data_all,
                               labels_train,
                               type = "uniform_fixed",
                               k_values = 5)
        
        acc = mean(isEqual(knn_out[names(labels_train),"resubstituted_labels"],
                           labels_train), na.rm = TRUE)
        return(acc)
      }, simplify = FALSE)
      
      res = rbind(res,
                  data.frame(
                    genes = rep(length(genes), length(embeddings_names)),
                    type = names(embeddings_names),
                    Accuracy = unlist(embeddings_accuracy),
                    Accuracy_resub = unlist(embeddings_resub_accuracy),
                    querySample = rep(querySample, length(embeddings_names)),
                    referenceSample = rep(referenceSample, length(embeddings_names)),
                    Sim = sim
                  ))
      
      # save the results as we go:
      saveRDS(res, file = resFile)
      saveRDS(simGenes, file = simGenesFile)
    }
  }
}

print(dim(res))
```

Examine results

```{r}
res2 = readRDS("../../output/MGD_StabMap_example/MGD_StabMap_example_res.Rds")

typesToKeep = c("StabMap",
                "PCA",
                "MultiMAP",
                "UINMF")

res2$type <- factor(res2$type, levels = typesToKeep)

g = ggplot(subset(res2, type %in% typesToKeep), aes(x = factor(genes), y = Accuracy)) + 
  theme_classic() +
  stat_summary(aes(width = 0.8, fill = type), geom = "bar", fun = "mean", position = "dodge", alpha = 0.5) +
  stat_summary(aes(group = type, colour = type), geom = "errorbar", fun.data = "mean_se", position = "dodge", width = 0.8) +
  xlab("Number of genes") +
  labs(fill = "") +
  labs(colour = "") +
  theme_classic() +
  geom_smooth(aes(group = type, colour = type), fill = NA, method = "loess") + 
  scale_colour_manual(values = method_colours, aesthetics = c("colour", "fill")) +
  theme(legend.position = "none") +
  NULL
print(g)

ggsave(g, file = "../../Figures/raw/MGD_accuracy_barplots.pdf", height = 6, width = 10)
```

Generate a network graph for illustration

```{r}
assay_list = mockMosaicData(names = c("Reference", "Query"),
                            ncells = c(50,50),
                            ngenes = list(1:10, 6:15))

mdt = mosaicDataTopology(assay_list)
V(mdt)$size = 60
V(mdt)$color = method_colours["StabMap"]
V(mdt)$label.color = "white"
V(mdt)$frame.color = "darkgrey"
V(mdt)$label.family = "sans"
plot(mdt)

pdf("../../Figures/raw/MGD_mdt.pdf", height = 4, width = 4)
par(mar = c(1,1,1,1))
print(plot(mdt))
dev.off()
```


Finish

```{r}
sessionInfo()
```