---
title: "StabMap simulation: Mouse Gastrulation Data"
author: "Shila Ghazanfar"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
       html_document:
                     toc: true
                     toc_float:
                           collapsed: false
                           smooth_scroll: false
                     code_folding: hide
                     fig_width: 10 
                     fig_height: 8
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE,
                      cache = FALSE, cache.lazy = FALSE)
```

```{r}
set.seed(2021)
```

Load scripts and packages.

```{r}
library(MouseGastrulationData)
library(scattermore)
library(ggpubr)
source("../scripts/initialise.R")
```

Helper plotting function:

```{r}
pl = function(embedding_UMAP, name, coltype = "type", facet = FALSE) {
  ind = sample(rownames(embedding_UMAP))
  df = data.frame(cell = ind,
                  U1 = embedding_UMAP[ind,1],
                  U2 = embedding_UMAP[ind,2],
                  type = factor(type[ind]),
                  ctype = ct_all[ind])
  df$col = df[,coltype]
  p = ggplot(df, aes(x = U1, y = U2, colour = col)) + 
    # geom_point(size = 0.3) + # replaced with scattermore
    geom_scattermore(pointsize = 2.3) +
    theme_classic() + 
    theme(legend.position = "none") +
    theme(axis.text = element_blank()) +
    theme(axis.ticks = element_blank()) +
    xlab("") +
    ylab("") +
    ggtitle(name)
  
  if (facet) {
    p <- p + facet_wrap(~type, nrow = 2, scales = "free")  +
      NULL
  }
  
  if (coltype == "ctype") {
    p <- p + scale_colour_manual(values = celltype_colours)
  }
  
  if (coltype == "type") {
    p <- p + scale_colour_manual(values = c("Query cells" = "red", "Reference cells" = "black"))
  }
  
  return(p)
}
```

Load data via the MouseGastrulationData Bioconductor package.

```{r}
mt = MouseGastrulationData::AtlasSampleMetadata
samples = mt[mt[, "stage"] == "E8.5", "sample"]

atlas = EmbryoAtlasData(type = "processed", samples = samples)
atlas <- logNormCounts(atlas)

celltype_colours = MouseGastrulationData::EmbryoCelltypeColours

ct_all = atlas$celltype
names(ct_all) <- colnames(atlas)

table(atlas$sample)
```

Build simulation set up. Treat three samples as the reference data, and 
treat the fourth sample as a query dataset with fewer features - selected
randomly among the entire set of features. Perform StabMap embedding, and 
compare against PCA, UINMF, and MultiMAP.

Assess performance by predicting cell types of the query dataset, using the 
reference dataset cells as the training set, and reporting the classification 
accuracy of the query dataset.

```{r}
nGenes_all = c(50, 100, 250, 500, 1000, 2000, 5000)

names(nGenes_all) <- paste0("Sim_", seq_len(length(nGenes_all)))

labels = "celltype"
assayNameReference = "logcounts"
assayNameQuery = "logcounts"
doUMAP = FALSE

resFile = "../../output/MGD_StabMap_example/MGD_StabMap_example_res.Rds"

if (!file.exists(resFile)) {
  res = NULL
} else {
  res = readRDS(resFile)
}

for (sim in names(nGenes_all)) {
  
  nGenes = nGenes_all[sim]
  
  if (nGenes <= 50) {
    nPCs = 20
  } else {
    nPCs = 50
  }
  
  for (querySample in samples) {
    
    for (referenceSample in setdiff(samples,querySample)) {
      
      print(nGenes)
      print(paste0("Query: ", querySample))
      print(paste0("Reference: ", referenceSample))
      
      # check has this already been done?
      if (any(res[,"Sim"] == sim & 
              res[,"querySample"] == querySample & 
              res[,"referenceSample"] == referenceSample)) next
      
      referenceSCE = atlas[,atlas$sample %in% referenceSample]
      
      # select HVGs from the referenceSCE and subset
      decomp <- modelGeneVar(referenceSCE)
      hvgs <- rownames(decomp)[decomp$mean > 0.05 & decomp$bio > 0]
      referenceSCE <- referenceSCE[hvgs,]
      
      # select random genes from among the HVGs for the query data
      genes = sample(hvgs)[seq_len(nGenes)]
      querySCE = atlas[genes, atlas$sample %in% querySample]
      
      # data type factor
      type = factor(ifelse(c(colnames(referenceSCE), colnames(querySCE)) %in% colnames(referenceSCE),
                           "Reference cells", "Query cells"), levels = c("Reference cells", "Query cells"))
      names(type) <- c(colnames(referenceSCE), colnames(querySCE))
      
      # input files as lists
      SCE_list = list(Reference = referenceSCE, Query = querySCE)
      assayNames = list(Reference = assayNameReference, Query = assayNameQuery)
      assay_list = mapply(assay, SCE_list, assayNames)
      counts_list = mapply(assay, SCE_list, "counts")
      labels_list = list(Reference = setNames(referenceSCE$celltype, colnames(referenceSCE)))
      
      g = mosaicDataUpSet(assay_list)
      g = mosaicDataUpSet(assay_list,
                          text.scale = c(3,1,2,1,3,3), 
                          point.size = 2.8, 
                          line.size = 1,
                          sets.x.label = "Cells",
                          mainbar.y.label = "Features",
                          main.bar.color = method_colours["StabMap"],
                          set_size.numbers_size = TRUE)
      print(g)
      pdf(file = paste0("../../Figures/raw/MGD_", 
                        sim, "_", 
                        querySample, "_",
                        referenceSample,
                        "_upset.pdf"), height = 4, width = 4)
      print(g)
      dev.off()
      
      # perform PCA
      PC_embedding = mapPCA(lapply(assay_list, as.matrix), nPCs = nPCs)
      dim(PC_embedding)
      colnames(PC_embedding) <- paste0("PC_", seq_len(ncol(PC_embedding)))
      
      # perform UINMF
      UINMF_embedding = UINMF_wrap(counts_list = counts_list)
      dim(UINMF_embedding)
      
      StabMap_embedding = stabMap(lapply(assay_list, as.matrix),
                                  reference_list = "Reference",
                                  ncomponentsReference = nPCs,
                                  projectAll = TRUE,
                                  scale.center = FALSE,
                                  scale.scale = FALSE,
                                  plot = FALSE
      )
      dim(StabMap_embedding)
      # StabMap_embedding <- reWeightEmbedding(StabMap_embedding, weights = c("Reference_PC" = 1, "Query_PC" = 1))
      
      # plot(StabMap_embedding[,1:2], col = ifelse(rownames(StabMap_embedding) %in% colnames(assay_list[["Reference"]]), "red", "blue"))
      # plot(PC_embedding[,1:2], col = ifelse(rownames(PC_embedding) %in% colnames(assay_list[["Reference"]]), "red", "blue"))

      MultiMAP_embedding = MultiMAP_wrap(assay_list, verbose = TRUE)
      dim(MultiMAP_embedding)
      
      embeddings_names = c(
        "PCA" = "PC_embedding",
        "UINMF" = "UINMF_embedding",
        "MultiMAP" = "MultiMAP_embedding",
        "StabMap" = "StabMap_embedding"
      )
      
      nobatch = c("MultiMAP_embedding", "UINMF_embedding")
      noumap = c("MultiMAP_embedding")
      
      cnames = unlist(lapply(assay_list, colnames))
      batchFactor_ref <- ifelse(cnames %in% colnames(referenceSCE),"Reference", "Query")
      batchFactor = as.character(interaction(batchFactor_ref, colData(atlas)[cnames, "sample"]))
      names(batchFactor) <- cnames
      
      sapply(embeddings_names, function(nm) {
        print(nm)
        # can include optional flags for certain embeddings to 
        # not be passed through a batch correction stage
        # e.g. from MultiMAP
        if (nm %in% nobatch) {
          # i.e. do nothing
          print(paste0("no batch correction for ", nm))
          assign(paste0(nm, "_corrected"), get(nm), envir = .GlobalEnv)
        } else {
          assign(paste0(nm, "_corrected_none"), get(nm),
                 envir = .GlobalEnv)
          assign(paste0(nm, "_corrected_MNN"), reducedMNN_batchFactor(get(nm), batchFactor),
                 envir = .GlobalEnv)
          assign(paste0(nm, "_corrected_Harmony"), Harmony_batchFactor(get(nm), batchFactor),
                 envir = .GlobalEnv)
          tryCatch(assign(paste0(nm, "_corrected_Seurat"), Seurat_batchFactor(get(nm), batchFactor),
                 envir = .GlobalEnv),
                 error = function(cond) assign(paste0(nm, "_corrected_Seurat"), NULL,
                 envir = .GlobalEnv)) 
        }
      }, simplify = FALSE)
      
      embeddings_names_corrected = c(nobatch, apply(expand.grid(setdiff(embeddings_names, nobatch),
                                                                paste0("_corrected_", c("none", "MNN", "Harmony", "Seurat"))), 1, paste0, collapse = ""))
      
      if (FALSE) {
        embeddings_corrected_UMAP = sapply(embeddings_names_corrected, function(nm) {
          if (nm %in% noumap) return(get(nm)) # i.e. do nothing
          calculateUMAP_rnames(get(paste0(nm)))
        }, simplify = FALSE)
        
        p_corrected = 
          wrap_plots(
            c(
              mapply(pl, embeddings_corrected_UMAP, names(embeddings_corrected_UMAP), SIMPLIFY = FALSE),
              mapply(pl, embeddings_corrected_UMAP, names(embeddings_corrected_UMAP), MoreArgs = list(coltype = "ctype"), SIMPLIFY = FALSE)
            )) +
          plot_layout(nrow = 2, ncol = length(embeddings_corrected_UMAP), byrow = TRUE)
        print(p_corrected)
        
        
        # Plot the UMAPs
        gList = mapply(pl, embeddings_corrected_UMAP, names(embeddings_corrected_UMAP), MoreArgs = list(coltype = "ctype", facet = TRUE), SIMPLIFY = FALSE)
        
        g = wrap_plots(gList) + plot_layout(nrow = 1)
        print(g)
        
        gList2 = mapply(pl, embeddings_corrected_UMAP, names(embeddings_corrected_UMAP),
                        MoreArgs = list(coltype = "type", facet = FALSE), SIMPLIFY = FALSE)
        g = wrap_plots(gList2) + plot_layout(nrow = 1)
        print(g)
        
        g = wrap_plots(c(lapply(gList,
                                function(x) x + 
                                  theme(strip.background = element_rect(colour=NA, fill=NA)) + 
                                  theme(plot.title = element_text(hjust = 0.5))
        ),
        lapply(gList2,
               function(x) x + ggtitle(NULL)))) + 
          plot_layout(nrow = 2, heights = c(2,1))
        
        print(g)
        ggsave(g, file = paste0("../../Figures/raw/MGD_", sim, "_", querySample,
                                "_", referenceSample,
                                "_umaps.pdf"), height = 6, width = 9)
        
        g_leg = as_ggplot(get_legend(gList[[1]] + theme(legend.position = "right")
                                     + guides(colour = guide_legend(title = "",
                                                                    ncol = 2,
                                                                    override.aes = list(size = 6.5)))))
        ggsave(g_leg, file = paste0("../../Figures/raw/MGD_celltype_umaps_leg.pdf"), height = 5, width = 4)
        
        g_leg_type = as_ggplot(get_legend(gList2[[1]] + theme(legend.position = "right")
                                          + guides(colour = guide_legend(title = "",
                                                                         ncol = 3,
                                                                         override.aes = list(size = 6)))))
        ggsave(g_leg_type, file = paste0("../../Figures/raw/MGD_type_umaps_leg.pdf"), height = 1, width = 4)
      }
      
      # save the embeddings as Rdata objects
      save(list = c(embeddings_names_corrected,
                    "PC_embedding",
                    "StabMap_embedding"),
                    # "embeddings_corrected_UMAP"),
           # save(PC_embedding, StabMap_embedding, PC_embedding_corrected,
           #      StabMap_embedding_corrected, UINMF_embedding, MultiMAP_embedding, 
           #      embeddings_corrected_UMAP,
           file = paste0("../../output/MGD_StabMap_example/MGD_StabMap_example_embeddings_",
                         sim, "_", querySample, "_", referenceSample,
                         ".RData"))
      
      # predict cell type labels using knn with k = 5
      referenceLabels = colData(referenceSCE)[,labels]
      names(referenceLabels) = colnames(referenceSCE)
      
      queryLabels = colData(querySCE)[,labels]
      names(queryLabels) = colnames(querySCE)
      
      # calculate accuracy of the query cells
      embeddings_accuracy = sapply(embeddings_names_corrected, function(nm) {
        if (is.null(get(nm))) return(NA)
        
        print(nm)
        
        # only use cells with labels
        data_all = get(paste0(nm))
        labels_train = referenceLabels[!is.na(referenceLabels) & names(referenceLabels) %in% rownames(data_all)]
        
        knn_out = embeddingKNN(data_all,
                               labels_train,
                               type = "uniform_fixed",
                               k_values = 5)
        
        acc = mean(isEqual(knn_out[names(queryLabels),"predicted_labels"], queryLabels), na.rm = TRUE)
        
        print(acc)
        
        return(acc)
      }, simplify = FALSE)
      
      print(embeddings_accuracy)
      
      # calculate the resubstitution accuracy
      embeddings_resub_accuracy = sapply(embeddings_names_corrected, function(nm) {
        if (is.null(get(nm))) return(NA)
        print(nm)
        
        # only use cells with labels
        data_all = get(paste0(nm))
        labels_train = referenceLabels[!is.na(referenceLabels) & names(referenceLabels) %in% rownames(data_all)]
        
        knn_out = embeddingKNN(data_all,
                               labels_train,
                               type = "uniform_fixed",
                               k_values = 5)
        
        acc = mean(isEqual(knn_out[names(labels_train),"resubstituted_labels"],
                           labels_train), na.rm = TRUE)
        return(acc)
      }, simplify = FALSE)
      
      res = rbind(res,
                  data.frame(
                    genes = rep(length(genes), length(embeddings_names_corrected)),
                    type = embeddings_names_corrected,
                    Accuracy = unlist(embeddings_accuracy),
                    Accuracy_resub = unlist(embeddings_resub_accuracy),
                    querySample = rep(querySample, length(embeddings_names_corrected)),
                    referenceSample = rep(referenceSample, length(embeddings_names_corrected)),
                    Sim = sim
                  ))
      
      print(res)
      
      # save the results as we go:
      saveRDS(res, file = resFile)
    }
  }
}

print(dim(res))
```

Examine results

```{r}
res2 = readRDS("../../output/MGD_StabMap_example/MGD_StabMap_example_res.Rds")

typesToKeep = c("StabMap_embedding_corrected_MNN",
                "PC_embedding_corrected_MNN",
                "MultiMAP_embedding",
                "UINMF_embedding")

res3 = subset(res2, type %in% typesToKeep)
res3$type <- factor(res3$type, levels = typesToKeep)

g = ggplot(res3, aes(x = factor(genes), y = Accuracy)) + 
  theme_classic() +
  stat_summary(aes(width = 0.8, fill = type), geom = "bar", fun = "mean", position = "dodge", alpha = 0.5) +
  stat_summary(aes(group = type, colour = type), geom = "errorbar", fun.data = "mean_se", position = "dodge", width = 0.8) +
  xlab("Number of genes") +
  ylab("Cell type accuracy") +
  labs(fill = "") +
  labs(colour = "") +
  theme_classic() +
  geom_smooth(aes(group = type, colour = type), fill = NA, method = "loess", position = position_dodge(width = 0.8)) + 
  geom_point(aes(group = type, colour = type), size = 0.5, position = position_dodge(width = 0.8)) +
  theme(axis.title = element_text(size = 30)) +
  theme(axis.text = element_text(size = 15)) +
  scale_colour_manual(values = method_colours, aesthetics = c("colour", "fill")) +
  theme(legend.position = "none") +
  NULL
print(g)

ggsave(g, file = "../../Figures/raw/MGD_accuracy_barplots.pdf", height = 5, width = 6)
```

Compare relative performance with choice of downstream integration method

```{r}
typesToKeep = c("PC_embedding_corrected_Harmony",
                "PC_embedding_corrected_MNN",
                "PC_embedding_corrected_none",
                "PC_embedding_corrected_Seurat",
                "StabMap_embedding_corrected_Harmony",
                "StabMap_embedding_corrected_MNN",
                "StabMap_embedding_corrected_none",
                "StabMap_embedding_corrected_Seurat")

res3 = subset(res2, type %in% typesToKeep)
res3$type <- factor(res3$type, levels = typesToKeep)

res3$method <- gsub("_.*", "", res3$type)
res3$integration <- gsub(".*_", "", res3$type)

res3$id <- paste0(res3$genes, "_", res3$querySample, "_", res3$referenceSample, "_",
                  res3$integration) 

tab_PC = reshape::cast(subset(res3, method == "PC"), formula = id ~ ., value = "Accuracy")
tab_StabMap = reshape::cast(subset(res3, method == "StabMap"), formula = id ~ ., value = "Accuracy")
tab_diff = tab_StabMap
tab_diff$diff = tab_StabMap$`(all)` - tab_PC$`(all)`
tab_diff$genes <- as.numeric(gsub("_.*", "", tab_diff$id))
tab_diff$integration <- gsub(".*_", "", tab_diff$id)
tab_diff$integration <- factor(as.character(tab_diff$integration),
                               levels = c("none", "Harmony", "MNN", "Seurat"))

maxh = max(abs(tab_diff$diff), na.rm = TRUE)

g = ggplot(tab_diff, aes(x = factor(genes), y = diff)) + 
  theme_classic() +
  stat_summary(aes(width = 0.8, fill = integration), geom = "bar", fun = "mean", position = "dodge", alpha = 0.5) +
  stat_summary(aes(group = integration, colour = integration), geom = "errorbar", fun.data = "mean_se", position = "dodge", width = 0.8) +
  xlab("Number of genes") +
  ylab("Cell type accuracy difference:\nStabMap - PCA") +
  labs(fill = "") +
  labs(colour = "") +
  theme_classic() +
  geom_smooth(aes(group = integration, colour = integration), fill = NA, method = "loess") + 
  geom_point(aes(group = integration, colour = integration), position = position_dodge(width = 0.8)) + 
  theme(axis.title = element_text(size = 20)) +
  theme(axis.text = element_text(size = 15)) +
  scale_colour_brewer(palette = "Set2", aesthetics = c("colour", "fill")) +
  # scale_colour_manual(values = method_colours, aesthetics = c("colour", "fill")) +
  theme(legend.position = "right") +
  geom_hline(yintercept = 0) +
  ylim(c(-1,1)*maxh) +
  NULL
print(g)

ggsave(g, file = "../../Figures/raw/MGD_relative_accuracy_PC_StabMap_barplots.pdf", height = 5, width = 6)
```


Generate a network graph for illustration

```{r}
assay_list = mockMosaicData(names = c("Reference", "Query"),
                            ncells = c(50,50),
                            ngenes = list(1:10, 6:15))

mdt = mosaicDataTopology(assay_list)
V(mdt)$size = 60
V(mdt)$color = method_colours["StabMap"]
V(mdt)$label.color = "white"
V(mdt)$frame.color = "darkgrey"
V(mdt)$label.family = "sans"
plot(mdt)

pdf("../../Figures/raw/MGD_mdt.pdf", height = 4, width = 4)
par(mar = c(1,1,1,1))
print(plot(mdt))
dev.off()
```

# Generate a UMAP graph for illustration

```{r}
sim = "Sim_3"
querySample = 29
referenceSample = 17
load(paste0("../../output/MGD_StabMap_example/MGD_StabMap_example_embeddings_",
            sim, "_", querySample, "_", referenceSample,
            ".RData"))
# names(embeddings_corrected_UMAP)
embeddings_corrected_UMAP = list(
  UINMF = calculateUMAP_rnames(UINMF_embedding),
  StabMap = calculateUMAP_rnames(StabMap_embedding_corrected_MNN),
  PCA = calculateUMAP_rnames(PC_embedding_corrected_MNN),
  MultiMAP = MultiMAP_embedding
)

referenceSCE = atlas[, atlas$sample %in% referenceSample]

querySCE = atlas[, atlas$sample %in% querySample]

# data type factor
type = factor(ifelse(c(colnames(referenceSCE), colnames(querySCE)) %in% colnames(referenceSCE),
                     "Reference cells", "Query cells"), levels = c("Reference cells", "Query cells"))
names(type) <- c(colnames(referenceSCE), colnames(querySCE))

p_corrected = 
  wrap_plots(
    c(
      mapply(pl, embeddings_corrected_UMAP, names(embeddings_corrected_UMAP), SIMPLIFY = FALSE),
      mapply(pl, embeddings_corrected_UMAP, names(embeddings_corrected_UMAP), MoreArgs = list(coltype = "ctype"), SIMPLIFY = FALSE)
    )) +
  plot_layout(nrow = 2, ncol = length(embeddings_corrected_UMAP), byrow = TRUE)
print(p_corrected)


# Plot the UMAPs
gList = mapply(pl, embeddings_corrected_UMAP, names(embeddings_corrected_UMAP), MoreArgs = list(coltype = "ctype", facet = TRUE), SIMPLIFY = FALSE)

g = wrap_plots(gList[c("StabMap", "PCA", "MultiMAP", "UINMF")]) + plot_layout(nrow = 1)
print(g)

gList2 = mapply(pl, embeddings_corrected_UMAP, names(embeddings_corrected_UMAP),
                MoreArgs = list(coltype = "type", facet = FALSE), SIMPLIFY = FALSE)
g = wrap_plots(gList2) + plot_layout(nrow = 1)
print(g)

g = wrap_plots(c(
  lapply(gList2[c("StabMap", "PCA", "UINMF", "MultiMAP")],
         function(x) x + 
           theme(plot.title = element_text(hjust = 0.5, size = 25)) + 
           NULL),
  lapply(gList[c("StabMap", "PCA", "UINMF", "MultiMAP")],
         function(x) x + 
           theme(strip.background = element_rect(colour=NA, fill=NA)) +
           theme(strip.text = element_text(size = 15)) +
           theme(plot.title = element_blank()) +
           NULL)
)) + 
  plot_layout(nrow = 2, heights = c(1,2))

print(g)
ggsave(g, file = paste0("../../Figures/raw/MGD_", sim, "_", querySample,
                        "_", referenceSample,
                        "_umaps.pdf"), height = 8, width = 9)

g_leg = as_ggplot(get_legend(gList[[1]] + theme(legend.position = "right")
                             + guides(colour = guide_legend(title = "",
                                                            ncol = 1,
                                                            override.aes = list(size = 6.5)))))
ggsave(g_leg, file = paste0("../../Figures/raw/MGD_celltype_umaps_leg.pdf"), height = 10, width = 4)

g_leg_type = as_ggplot(get_legend(gList2[[1]] + theme(legend.position = "right")
                                  + guides(colour = guide_legend(title = "",
                                                                 ncol = 3,
                                                                 override.aes = list(size = 6)))))
ggsave(g_leg_type, file = paste0("../../Figures/raw/MGD_type_umaps_leg.pdf"), height = 1, width = 4)
```

# Generate a UMAP graph for illustration

```{r}
sims = c("Sim_1", "Sim_2", "Sim_4", "Sim_5", "Sim_6", "Sim_7")
for (sim in sims) {
  querySample = 29
  referenceSample = 17
  load(paste0("../../output/MGD_StabMap_example/MGD_StabMap_example_embeddings_",
              sim, "_", querySample, "_", referenceSample,
              ".RData"))
  embeddings_corrected_UMAP = list(
    UINMF = calculateUMAP_rnames(UINMF_embedding),
    StabMap = calculateUMAP_rnames(StabMap_embedding_corrected_MNN),
    PCA = calculateUMAP_rnames(PC_embedding_corrected_MNN),
    MultiMAP = MultiMAP_embedding
  )
  
  # genes = simGenes[[sim]]
  
  referenceSCE = atlas[,atlas$sample %in% referenceSample]
  
  querySCE = atlas[seq_len(nGenes_all[sim]),atlas$sample %in% querySample]
  
  # data type factor
  type = factor(ifelse(c(colnames(referenceSCE), colnames(querySCE)) %in% colnames(referenceSCE),
                       "Reference cells", "Query cells"), levels = c("Reference cells", "Query cells"))
  names(type) <- c(colnames(referenceSCE), colnames(querySCE))
  
  p_corrected = 
    wrap_plots(
      c(
        mapply(pl, embeddings_corrected_UMAP, names(embeddings_corrected_UMAP), SIMPLIFY = FALSE),
        mapply(pl, embeddings_corrected_UMAP, names(embeddings_corrected_UMAP), MoreArgs = list(coltype = "ctype"), SIMPLIFY = FALSE)
      )) +
    plot_layout(nrow = 2, ncol = length(embeddings_corrected_UMAP), byrow = TRUE)
  print(p_corrected)
  
  
  # Plot the UMAPs
  gList = mapply(pl, embeddings_corrected_UMAP, names(embeddings_corrected_UMAP), MoreArgs = list(coltype = "ctype", facet = TRUE), SIMPLIFY = FALSE)
  
  g = wrap_plots(gList[c("StabMap", "PCA", "MultiMAP", "UINMF")]) + plot_layout(nrow = 1)
  print(g)
  
  gList2 = mapply(pl, embeddings_corrected_UMAP, names(embeddings_corrected_UMAP),
                  MoreArgs = list(coltype = "type", facet = FALSE), SIMPLIFY = FALSE)
  g = wrap_plots(gList2) + plot_layout(nrow = 1)
  print(g)
  
  g = wrap_plots(c(
    lapply(gList2[c("StabMap", "PCA", "UINMF", "MultiMAP")],
           function(x) x + 
             theme(plot.title = element_text(hjust = 0.5, size = 25)) + 
             NULL),
    lapply(gList[c("StabMap", "PCA", "UINMF", "MultiMAP")],
           function(x) x + 
             theme(strip.background = element_rect(colour=NA, fill=NA)) +
             theme(strip.text = element_text(size = 15)) +
             theme(plot.title = element_blank()) +
             NULL)
  )) + 
    plot_layout(nrow = 2, heights = c(1,2))
  
  print(g)
  ggsave(g, file = paste0("../../Figures/raw/MGD_", sim, "_", querySample,
                          "_", referenceSample,
                          "_umaps.pdf"), height = 8, width = 9)
  
  assay_list = list(Reference = logcounts(referenceSCE),
                    Query = logcounts(querySCE))
  
  g = mosaicDataUpSet(assay_list,
                      text.scale = c(3,1,2,1,3,3), 
                      point.size = 2.8, 
                      line.size = 1,
                      sets.x.label = "Cells",
                      mainbar.y.label = "Features",
                      main.bar.color = method_colours["StabMap"],
                      set_size.numbers_size = TRUE,
                      mainbar.y.max = max(unlist(lapply(assay_list, nrow))*1.2))
  print(g)
  pdf(file = paste0("../../Figures/raw/MGD_", 
                    sim, "_", 
                    querySample, "_",
                    referenceSample,
                    "_upset.pdf"), height = 4, width = 4)
  print(g)
  dev.off()
}
```


Finish

```{r}
sessionInfo()
```
