---
title: "StabMap simulation: Mouse gastrulation mosaic integration with missing cell types"
author: "Shila Ghazanfar"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
       html_document:
                     toc: true
                     toc_float:
                           collapsed: false
                           smooth_scroll: false
                     code_folding: hide
                     fig_width: 10 
                     fig_height: 8
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE,
                      cache = FALSE, cache.lazy = FALSE)
```

```{r}
set.seed(2021)
```

Load scripts and packages.

```{r}
library(MouseGastrulationData)
library(scattermore)
library(ggpubr)
library(dplyr)
source("../scripts/initialise.R")
```

Helper plotting function:

```{r}
pl = function(embedding_UMAP, name, coltype = "type", facet = FALSE) {
  ind = sample(rownames(embedding_UMAP))
  df = data.frame(cell = ind,
                  U1 = embedding_UMAP[ind,1],
                  U2 = embedding_UMAP[ind,2],
                  type = factor(type[ind]),
                  ctype = ct_all[ind])
  df$col = df[,coltype]
  p = ggplot(df, aes(x = U1, y = U2, colour = col)) + 
    geom_point(size = 0.3) + # replaced with scattermore
    # geom_scattermore(pointsize = 2.3) +
    theme_classic() + 
    theme(legend.position = "none") +
    theme(axis.text = element_blank()) +
    theme(axis.ticks = element_blank()) +
    xlab("") +
    ylab("") +
    ggtitle(name)
  
  if (facet) {
    p <- p + facet_wrap(~type, nrow = 2, scales = "free")  +
      NULL
  }
  
  if (coltype == "ctype") {
    p <- p + scale_colour_manual(values = celltype_colours)
  }
  
  if (coltype == "type") {
    # p <- p + scale_colour_manual(values = c("Query cells" = "red", "Reference cells" = "black"))
    p <- p + scale_colour_discrete()
  }
  
  return(p)
}
```

Load data via the MouseGastrulationData Bioconductor package.

```{r}
mt = MouseGastrulationData::AtlasSampleMetadata
samples = mt[mt[, "stage"] == "E8.5", "sample"]

atlas = EmbryoAtlasData(type = "processed", samples = samples)
atlas <- atlas[,!is.na(atlas$celltype)]
atlas <- logNormCounts(atlas)

decomp <- modelGeneVar(atlas)
hvgs <- rownames(decomp)[decomp$mean > 0.05 & decomp$total > 0]
length(hvgs)
atlas <- atlas[hvgs,]

celltype_colours = MouseGastrulationData::EmbryoCelltypeColours

ct_all = atlas$celltype
names(ct_all) <- colnames(atlas)

table(atlas$sample)
```

Build simulation set up. Take the dataset and split into 3 groups.
Then select 3000 features from the HVGs and with 50% percent overlap
between contiguous datasets. Remove various cell types from the bridge dataset. 
Then jointly embed the data using StabMap, and record the cell type 
classification accuracy.

```{r}
k = 3
nGenes = 2000
nCells = 3000
labels = "celltype"
celltypesToRemove = c("None", sort(names(table(atlas$celltype)[table(atlas$celltype) >= 100])))

acc_prop_df_all = NULL

for (i in 1:5) {
  
  print(i)
  print(nCells)
  print(nGenes)
  
  # check there are enough features, this should be FALSE
  ceiling((k+1)*(nGenes/2)) > length(hvgs)
  
  features_to_split = sample(hvgs, ceiling((k+1)*(nGenes/2)))
  feature_split_all = split(features_to_split, rep(seq_len(k+1), length.out = length(features_to_split)))
  feature_split = list()
  for (kk in 1:k) {
    feature_split[[kk]] <- unlist(feature_split_all[kk:(kk+1)])
  }
  names(feature_split) <- paste0("Dataset_",seq_len(k))
  
  atlas_cells = sample(colnames(atlas))[seq_len(nCells*k)]
  cell_split_orig = split(atlas_cells, rep(seq_len(k), length.out = length(atlas_cells)))
  
  for (celltypeToRemove in celltypesToRemove) {
    
    print(celltypeToRemove)
    
    cell_split = cell_split_orig
    
    # Remove celltype from the middle dataset
    ct_cells = colnames(atlas)[atlas$celltype == celltypeToRemove]
    cell_split[[2]] <- setdiff(cell_split[[2]], ct_cells)
    
    assay_list = mapply(function(feat,cel) {
      logcounts(atlas)[feat,cel]
    }, feature_split, cell_split, SIMPLIFY = FALSE)
    lapply(assay_list, dim)
    plot(mosaicDataTopology(assay_list))
    
    type = setNames(rep(paste0("Dataset_",names(cell_split)), times = unlist(lapply(cell_split, length))), unlist(cell_split))
    
    g = mosaicDataUpSet(assay_list,
                        text.scale = c(3,1,2,1,3,3), 
                        point.size = 2.8, 
                        line.size = 1,
                        sets.x.label = "Cells",
                        mainbar.y.label = "Features",
                        main.bar.color = method_colours["StabMap"],
                        set_size.numbers_size = TRUE)
    print(g)
    # pdf(file = paste0("../../Figures/raw/MGD_", 
    #                   sim, "_", 
    #                   querySample, "_",
    #                   referenceSample,
    #                   "_upset.pdf"), height = 4, width = 4)
    # print(g)
    # dev.off()
    
    nPCs = ifelse(nGenes <= 50, 10, 50)
    
    StabMap_embedding = stabMap(assay_list,
                                reference_list = "Dataset_2",
                                ncomponentsReference = nPCs,
                                ncomponentsSubset = nPCs)
    dim(StabMap_embedding)
    
    StabMap_embedding_all = stabMap(assay_list,
                                    ncomponentsReference = nPCs,
                                    ncomponentsSubset = nPCs)
    dim(StabMap_embedding_all)
    StabMap_embedding_all <- reWeightEmbedding(StabMap_embedding_all)
    
    # now perform cell type classification using the embedding
    # predict cell type labels using knn with k = 5
    referenceLabels = colData(atlas)[colnames(assay_list[["Dataset_1"]]),labels]
    names(referenceLabels) = colnames(assay_list[["Dataset_1"]])
    
    queryLabels = colData(atlas)[colnames(assay_list[["Dataset_3"]]),labels]
    names(queryLabels) = colnames(assay_list[["Dataset_3"]])
    
    # calculate accuracy of the query cells
    labels_train = referenceLabels
    
    data = StabMap_embedding
    
    knn_out = embeddingKNN(data,
                           labels_train,
                           type = "uniform_fixed",
                           k_values = 5)
    
    acc = isEqual(knn_out[names(queryLabels),"predicted_labels"], queryLabels)
    
    # prop.acc = prop.table(table(acc, type[names(queryLabels)]), 2)["1",]
    # prop.acc[gtools::mixedorder(names(prop.acc))]
    # barplot(prop.acc[gtools::mixedorder(names(prop.acc))][-1])
    acc_prop = tapply(acc, queryLabels, mean)
    
    acc_prop_df = data.frame(celltypeToRemove = celltypeToRemove,
                             celltype = names(acc_prop),
                             acc = acc_prop
                             # method = "Dataset_2"
    )
    
    # acc_prop_df_all <- rbind(acc_prop_df_all, acc_prop_df)
    
    # Now for the alternative stabmap output
    
    data = StabMap_embedding_all
    
    knn_out = embeddingKNN(data,
                           labels_train,
                           type = "uniform_fixed",
                           k_values = 5)
    
    acc = isEqual(knn_out[names(queryLabels),"predicted_labels"], queryLabels)
    
    # prop.acc = prop.table(table(acc, type[names(queryLabels)]), 2)["1",]
    # prop.acc[gtools::mixedorder(names(prop.acc))]
    # barplot(prop.acc[gtools::mixedorder(names(prop.acc))][-1])
    acc_prop = tapply(acc, queryLabels, mean)
    
    acc_prop_df$acc_all = acc_prop
    # acc_prop_df = data.frame(celltypeToRemove = celltypeToRemove,
    #                          celltype = names(acc_prop),
    #                          acc = acc_prop,
    #                          method = "All")
    
    acc_prop_df_all <- rbind(acc_prop_df_all, acc_prop_df)
    
  }
}
```

Save results table

```{r}
saveRDS(acc_prop_df_all, file = "../../output/MGD_MissingCellTypes_df.Rds")
```

Generate plot

```{r}
# library(shades)
acc_prop_df_all$acc_diff = acc_prop_df_all$acc_all - acc_prop_df_all$acc

acc_prop_df_all_means = acc_prop_df_all %>%
  group_by(celltypeToRemove, celltype) %>%
  dplyr::summarise(acc_diff = mean(acc_diff))

g = ggplot(subset(acc_prop_df_all_means, celltype %in% celltypesToRemove), aes(x = celltype, y = celltypeToRemove)) + 
  geom_tile(aes(fill = acc_diff)) + # positive means better accuracy when using all reference
  # geom_tile(aes(fill = acc)) + # 
  # facet_wrap(~method) +
  # geom_line(aes(colour = factor(nGenes), group = rep), alpha = 0.1) + 
  # geom_point(aes(colour = factor(nGenes), group = rep), alpha = 0.1) + 
  # geom_smooth(aes(fill = factor(nGenes), group = factor(nGenes), colour = factor(nGenes)), alpha = 0.3) +
  # facet_wrap(~nCells, nrow = 1, labeller = labeller(nCells = function(x) paste0(x, " cells per dataset"))) + 
  # ylab("Accuracy") +
  # xlab("Dataset") +
  theme_classic() + 
  coord_fixed() +
  xlab("") +
  ylab("Removed cell type") +
  theme(legend.position = "none") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1)) +
  scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0) +
  guides(fill = guide_colourbar(title = "Mean difference\nin accuracy")) +
  # lightness(scale_colour_brewer(palette = 1, aesthetics = c("colour", "fill")), scalefac(0.7)) +
  NULL

g

ggsave(g, file = "../../Figures/raw/MGD_MissingCellType_accuracy_heatmap.pdf",
       height = 8, width = 8)

g_leg = as_ggplot(get_legend(g + theme(legend.position = "right")))

g_leg
ggsave(g_leg, file = "../../Figures/raw/MGD_MissingCellType_accuracy_heatmap_leg.pdf",
       height = 8, width = 6)

g_leg2 = as_ggplot(get_legend(g + theme(legend.position = "bottom")))

g_leg2
ggsave(g_leg2, file = "../../Figures/raw/MGD_MissingCellType_accuracy_heatmap_leg2.pdf",
       height = 3, width = 12)


# scatterplot per cell type of accuracy with and without removing the ct
celltypesToRemoveSub <- sort(setdiff(celltypesToRemove, "None"))
acc_prop_df_all_sub = subset(acc_prop_df_all, celltypeToRemove == "None")

ct_acc_df = data.frame(
  celltype = celltypesToRemoveSub,
  acc = tapply(acc_prop_df_all_sub$acc, acc_prop_df_all_sub$celltype, mean)[celltypesToRemoveSub],
  acc_all = tapply(acc_prop_df_all_sub$acc_all, acc_prop_df_all_sub$celltype, mean)[celltypesToRemoveSub]
)

ct_acc_df$acc_rm = as.numeric(sapply(ct_acc_df$celltype, function(ct) {
  mean(subset(acc_prop_df_all, celltypeToRemove == ct & celltype == ct)[,"acc"])
}, simplify = TRUE))
ct_acc_df$acc_rm_all = as.numeric(sapply(ct_acc_df$celltype, function(ct) {
  mean(subset(acc_prop_df_all, celltypeToRemove == ct & celltype == ct)[,"acc_all"])
}, simplify = TRUE))

g = ggplot(subset(ct_acc_df, celltype %in% celltypesToRemove)) + 
  geom_point(aes(x = acc, y = acc_rm, colour = celltype, shape = "Dataset 2")) + 
  geom_smooth(aes(x = acc, y = acc_rm), method = "lm", colour = "grey", fill = NA, alpha = 0.5) +
  
  geom_point(aes(x = acc_all, y = acc_rm_all, colour = celltype, shape = "All")) + 
  geom_smooth(aes(x = acc_all, y = acc_rm_all), method = "lm", colour = "blue", fill = NA, alpha = 0.5) +
  
  geom_abline(slope = 1, intercept = 0, linetype = "dotted") + 
  scale_colour_manual(values = celltype_colours[celltypesToRemoveSub]) +
  scale_shape_manual(values = c("Dataset 2" = 16, "All" = 17)) +
  scale_fill_manual(values = c("Dataset 2" = "grey", "All" = "blue")) +
  theme_classic() +
  xlab("Accuracy including cell type") +
  ylab("Accuracy removing cell type") +
  xlim(c(0,1)) +
  ylim(c(0,1)) +
  theme(legend.position = "none") +
  guides(colour = guide_legend(title = "")) + 
  guides(shape = guide_legend(title = "", override.aes = list(colour = c("grey", "blue")))) +
  NULL

g

ggsave(g, file = "../../Figures/raw/MGD_MissingCellType_accuracy_scatterplot.pdf",
       height = 6, width = 6)

g_leg = as_ggplot(get_legend(g + theme(legend.position = "right")))

g_leg
ggsave(g_leg, file = "../../Figures/raw/MGD_MissingCellType_accuracy_scatterplot_leg.pdf",
       height = 8, width = 4)

```


Finish

```{r}
sessionInfo()
```