---
title: "StabMap simulation: Mouse gastrulation data mosaic integration with Multihop data"
author: "Shila Ghazanfar"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
       html_document:
                     toc: true
                     toc_float:
                           collapsed: false
                           smooth_scroll: false
                     code_folding: hide
                     fig_width: 10 
                     fig_height: 8
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE,
                      cache = FALSE, cache.lazy = FALSE)
```

```{r}
set.seed(2021)
```

Load scripts and packages.

```{r}
library(MouseGastrulationData)
library(scattermore)
library(ggpubr)
source("../scripts/initialise.R")
```

Helper plotting function:

```{r}
pl = function(embedding_UMAP, name, coltype = "type", facet = FALSE) {
  ind = sample(rownames(embedding_UMAP))
  df = data.frame(cell = ind,
                  U1 = embedding_UMAP[ind,1],
                  U2 = embedding_UMAP[ind,2],
                  type = factor(type[ind]),
                  ctype = ct_all[ind])
  df$col = df[,coltype]
  p = ggplot(df, aes(x = U1, y = U2, colour = col)) + 
    geom_point(size = 0.3) + # replaced with scattermore
    # geom_scattermore(pointsize = 2.3) +
    theme_classic() + 
    theme(legend.position = "none") +
    theme(axis.text = element_blank()) +
    theme(axis.ticks = element_blank()) +
    xlab("") +
    ylab("") +
    ggtitle(name)
  
  if (facet) {
    p <- p + facet_wrap(~type, nrow = 2, scales = "free")  +
      NULL
  }
  
  if (coltype == "ctype") {
    p <- p + scale_colour_manual(values = celltype_colours)
  }
  
  if (coltype == "type") {
    # p <- p + scale_colour_manual(values = c("Query cells" = "red", "Reference cells" = "black"))
    p <- p + scale_colour_discrete()
  }
  
  return(p)
}
```

Load data via the MouseGastrulationData Bioconductor package.

```{r}
mt = MouseGastrulationData::AtlasSampleMetadata
samples = mt[mt[, "stage"] == "E8.5", "sample"]

atlas = EmbryoAtlasData(type = "processed", samples = samples)
atlas <- logNormCounts(atlas)

decomp <- modelGeneVar(atlas)
hvgs <- rownames(decomp)[decomp$mean > 0.05 & decomp$total > 0]
length(hvgs)
atlas <- atlas[hvgs,]

celltype_colours = MouseGastrulationData::EmbryoCelltypeColours

ct_all = atlas$celltype
names(ct_all) <- colnames(atlas)

table(atlas$sample)
```

Build simulation set up. Take the dataset and split into k groups.
Then select up to G/k features from the HVGs and determine the p percent overlap
between contiguous datasets. Then jointly embed the data using StabMap, and 
record the cell type classification accuracy.

```{r}
k = 8
nGenes_all = c(100, 200, 500, 1000)
nCells_all = c(500, 1000, 2000)
labels = "celltype"

acc_df_all = NULL

for (i in 1:5) {
  
  for (nCells in nCells_all) {
    
    for (nGenes in nGenes_all) {
      
      print(i)
      print(nCells)
      print(nGenes)
      
      # check there are enough features, this should be FALSE
      ceiling((k+1)*(nGenes/2)) > length(hvgs)
      
      features_to_split = sample(hvgs, ceiling((k+1)*(nGenes/2)))
      feature_split_all = split(features_to_split, rep(seq_len(k+1), length.out = length(features_to_split)))
      feature_split = list()
      for (kk in 1:k) {
        feature_split[[kk]] <- unlist(feature_split_all[kk:(kk+1)])
      }
      names(feature_split) <- paste0("Dataset_",seq_len(k))
      
      atlas_cells = sample(colnames(atlas))[seq_len(nCells*8)]
      cell_split = split(atlas_cells, rep(seq_len(k), length.out = length(atlas_cells)))
      
      assay_list = mapply(function(feat,cel) {
        logcounts(atlas)[feat,cel]
      }, feature_split, cell_split, SIMPLIFY = FALSE)
      lapply(assay_list, dim)
      plot(mosaicDataTopology(assay_list))
      
      type = setNames(rep(paste0("Dataset_",names(cell_split)), times = unlist(lapply(cell_split, length))), unlist(cell_split))
      
      g = mosaicDataUpSet(assay_list,
                          text.scale = c(3,1,2,1,3,3), 
                          point.size = 2.8, 
                          line.size = 1,
                          sets.x.label = "Cells",
                          mainbar.y.label = "Features",
                          main.bar.color = method_colours["StabMap"],
                          set_size.numbers_size = TRUE)
      print(g)
      # pdf(file = paste0("../../Figures/raw/MGD_", 
      #                   sim, "_", 
      #                   querySample, "_",
      #                   referenceSample,
      #                   "_upset.pdf"), height = 4, width = 4)
      # print(g)
      # dev.off()
      
      nPCs = ifelse(nGenes <=50, 10, 50)
      
      StabMap_embedding = stabMap(assay_list,
                                  reference_list = "Dataset_1",
                                  ncomponentsReference = nPCs,
                                  ncomponentsSubset = nPCs,
                                  projectAll = TRUE)
      dim(StabMap_embedding)
      
      # now perform cell type classification using the embedding
      # predict cell type labels using knn with k = 5
      referenceLabels = colData(atlas)[colnames(assay_list[["Dataset_1"]]),labels]
      names(referenceLabels) = colnames(assay_list[["Dataset_1"]])
      
      queryLabels = colData(atlas)[,labels]
      names(queryLabels) = colnames(atlas)
      
      # calculate accuracy of the query cells
      # only use cells with labels
      data_all = StabMap_embedding
      labels_train = referenceLabels[!is.na(referenceLabels) & names(referenceLabels) %in% rownames(data_all)]
      
      knn_out = embeddingKNN(data_all,
                             labels_train,
                             type = "uniform_fixed",
                             k_values = 5)
      
      acc = isEqual(knn_out[names(queryLabels),"predicted_labels"], queryLabels)
      
      prop.acc = prop.table(table(acc, type[names(queryLabels)]), 2)["1",]
      prop.acc[gtools::mixedorder(names(prop.acc))]
      barplot(prop.acc[gtools::mixedorder(names(prop.acc))][-1])
      
      acc_df = data.frame(nCells = nCells,
                          nGenes = nGenes,
                          acc = prop.acc,
                          Dataset = names(prop.acc),
                          rep = paste0(c(nCells,nGenes,i), collapse = "_"))
      
      acc_df_all <- rbind(acc_df_all, acc_df)
      
    }
    
  }
  
}
```

Save results table

```{r}
saveRDS(acc_df_all, file = "../../output/MGD_multihop_example_df.Rds")
```

Generate plot

```{r}
library(shades)
g = ggplot(subset(acc_df_all, Dataset != "Dataset_1"), aes(x = Dataset, y = acc)) + 
  geom_line(aes(colour = factor(nGenes), group = rep), alpha = 0.1) + 
  geom_point(aes(colour = factor(nGenes), group = rep), alpha = 0.1) + 
  geom_smooth(aes(fill = factor(nGenes), group = factor(nGenes), colour = factor(nGenes)), alpha = 0.3) +
  facet_wrap(~nCells, nrow = 1, labeller = labeller(nCells = function(x) paste0(x, " cells per dataset"))) + 
  ylab("Accuracy") +
  xlab("Dataset") +
  theme_classic() + 
  theme(legend.position = "none") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1)) + 
  lightness(scale_colour_brewer(palette = 1, aesthetics = c("colour", "fill")), scalefac(0.7)) +
  NULL

g

ggsave(g, file = "../../Figures/raw/MGD_multihop_accuracy.pdf",
       height = 5, width = 12)

g_leg = as_ggplot(get_legend(g + theme(legend.position = "right") + guides(colour = guide_legend(title = "Number of genes \n per dataset"),
                                                                           fill = guide_legend(title = "Number of genes \n per dataset"))))

g_leg
ggsave(g_leg, file = "../../Figures/raw/MGD_multihop_accuracy_leg.pdf",
       height = 8, width = 6)

g_leg2 = as_ggplot(get_legend(g + theme(legend.position = "bottom") + guides(colour = guide_legend(title = "Number of genes \n per dataset"),
                                                                           fill = guide_legend(title = "Number of genes \n per dataset"))))

g_leg2
ggsave(g_leg2, file = "../../Figures/raw/MGD_multihop_accuracy_leg2.pdf",
       height = 3, width = 12)

```


Finish

```{r}
sessionInfo()
```