---
title: "StabMap simulation: Mouse Gastrulation Data"
author: "Shila Ghazanfar"
date: "14/05/2021"
output:
       html_document:
                     toc: true
                     toc_float:
                           collapsed: false
                           smooth_scroll: false
                     code_folding: hide
                     fig_width: 10 
                     fig_height: 8
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE,
                      cache = TRUE,cache.lazy = FALSE)
```

```{r}
set.seed(2021)
```

Load scripts and packages.

```{r}
library(MouseGastrulationData)
library(scater)
library(ggplot2)
source("../scripts/adaptiveKNN.R")
source("../scripts/StabMap_functions.R")
```

Load data via the MouseGastrulationData Bioconductor package

```{r}
mt = MouseGastrulationData::AtlasSampleMetadata
samples = mt[mt[, "stage"] == "E8.5", "sample"]

atlas = EmbryoAtlasData(type = "processed", samples = samples)
atlas <- logNormCounts(atlas)

celltype_colours = MouseGastrulationData::EmbryoCelltypeColours

ct_all = atlas$celltype
names(ct_all) <- colnames(atlas)

# the sample information is used in the simulation below
table(atlas$sample)
```

Build simulation set up. Treat three samples as the reference data, and 
treat the fourth sample as a query dataset with fewer features - selected
randomly among the entire set of features. Perform StabMap embedding, both
ignoring and using cell type labels, and compare against PCA, UINMF, and
MultiMAP.

Assess performance by predicting cell types of the query dataset, using the 
reference dataset cells as the training set, and reporting the classification 
accuracy of the query dataset.

Option to plot the UMAPs for each simulation, currently set to not run.

```{r}
nGenes_all = rep(rev(c(100, 150, 200, 250, 500, 1000, 2500, 5000, 10000)),
                 times = 3)
names(nGenes_all) <- paste0("Sim_", seq_len(length(nGenes_all)))

labels = "celltype"
assayNameReference = "logcounts"
assayNameQuery = "logcounts"

res = NULL
simGenes = list()

for (sim in names(nGenes_all)) {
  
  nGenes = nGenes_all[sim]
  genes = sample(rownames(atlas)[rowVars(assay(atlas, assayNameReference)) > 0])[seq_len(nGenes)]
  simGenes[[sim]] <- genes
  
  for (querySample in samples) {
    
    print(nGenes)
    print(querySample)
    
    doUMAP = FALSE
    
    # build reference and query data using querySample and genes
    referenceSCE = atlas[,atlas$sample != querySample]
    
    querySCE = atlas[genes, atlas$sample == querySample]
    
    # data type factor
    type = factor(ifelse(c(colnames(referenceSCE), colnames(querySCE)) %in% colnames(referenceSCE),
                         "reference", "query"))
    names(type) <- c(colnames(referenceSCE), colnames(querySCE))
    
    # input files as lists
    SCE_list = list(reference = referenceSCE, query = querySCE)
    assayNames = list(reference = assayNameReference, query = assayNameQuery)
    assay_list = mapply(assay, SCE_list, assayNames)
    counts_list = mapply(assay, SCE_list, "counts")
    
    # perform PCA
    PC_embedding = mapPCA(assay_list)
    dim(PC_embedding)
    
    # perform UINMF
    UINMF_embedding = UINMF_wrap(counts_list = counts_list)
    dim(UINMF_embedding)
    
    # Perform linear discriminant stabMap
    LD_embedding = stabMapLabelled(referenceSCE = referenceSCE,
                                   query_assay = assay_list[["query"]],
                                   labels = labels,
                                   assayNameReference = assayNameReference,
                                   prop_explained = 1,
                                   selectBestLDA = FALSE)
    dim(LD_embedding)
    
    # comparative stabmap
    SC_embedding = stabMapComparative(assay_list)
    dim(SC_embedding)
    
    # stabMap treating larger data as only reference
    SC_ref_embedding = stabMapComparative(assay_list,
                                          stabilise = c(TRUE, FALSE))
    dim(SC_ref_embedding)
    
    # LDA and comparative stabmap together
    LD_SC_embedding = cbind(LD_embedding, SC_embedding)
    
    # placeholder for MultiMAP, make sure it's added to the below
    # embeddings_names object
    
    embeddings_names = c("StabMap" = "SC_embedding",
                         "StabMap_LD" = "LD_SC_embedding",
                         "PCA" = "PC_embedding",
                         "UINMF" = "UINMF_embedding",
                         "StabMap_ref" = "SC_ref_embedding",
                         "StabMap_LD_only" = "LD_embedding"
                         )
    
    if (doUMAP) {
      
      embeddings_UMAP = sapply(embeddings_names, function(nm){
        calculateUMAP_rnames(get(nm))
      }, simplify = FALSE)
      
      pl = function(embedding_UMAP, name, coltype = "type") {
        ind = sample(rownames(embedding_UMAP))
        df = data.frame(cell = ind,
                        U1 = embedding_UMAP[ind,1],
                        U2 = embedding_UMAP[ind,2],
                        type = factor(type[ind]),
                        ctype = ct_all[ind])
        df$col = df[,coltype]
        p = ggplot(df, aes(x = U1, y = U2, colour = col)) + 
          geom_point() +
          theme_classic() + 
          theme(legend.position = "none") +
          coord_fixed() +
          ggtitle(name)
        
        if (coltype == "ctype") {
          p <- p + scale_colour_manual(values = celltype_colours)
        }
        
        return(p)
      }
      
      p_uncorrected = 
        wrap_plots(
          c(
            mapply(pl, embeddings_UMAP, names(embeddings_UMAP), SIMPLIFY = FALSE),
            mapply(pl, embeddings_UMAP, names(embeddings_UMAP), MoreArgs = list(coltype = "ctype"), SIMPLIFY = FALSE)
          )) +
        plot_layout(nrow = 2, ncol = 6, byrow = TRUE)
      p_uncorrected
      
    }
    
    cnames = unlist(lapply(assay_list, colnames))
    batchFactor_ref <- ifelse(cnames %in% colnames(referenceSCE),"Reference", "Query")
    batchFactor = as.character(interaction(batchFactor_ref, colData(atlas)[cnames, "sample"]))
    names(batchFactor) <- cnames
    
    sapply(embeddings_names, function(nm) {
      print(nm)
      # can include optional flags for certain embeddings to 
      # not be passed through a batch correction stage
      # e.g. from MultiMAP
      assign(paste0(nm, "_corrected"), reducedMNN_batchFactor(get(nm), batchFactor),
             envir = .GlobalEnv)
    }, simplify = FALSE)
   
    if (doUMAP) {
    
      embeddings_corrected_UMAP = sapply(embeddings_names, function(nm){
        calculateUMAP_rnames(get(paste0(nm, "_corrected")))
      }, simplify = FALSE)
      
      p_corrected = 
        wrap_plots(
          c(
            mapply(pl, embeddings_corrected_UMAP, names(embeddings_corrected_UMAP), SIMPLIFY = FALSE),
            mapply(pl, embeddings_corrected_UMAP, names(embeddings_corrected_UMAP), MoreArgs = list(coltype = "ctype"), SIMPLIFY = FALSE)
          )) +
        plot_layout(nrow = 2, ncol = 6, byrow = TRUE)
      p_corrected
      
    }
    
    # predict cell type labels using knn with k = 5
    referenceLabels = colData(referenceSCE)[,labels]
    names(referenceLabels) = colnames(referenceSCE)
    
    queryLabels = colData(querySCE)[,labels]
    names(queryLabels) = colnames(querySCE)
    
    # calculate accuracy of the query cells
    embeddings_accuracy = sapply(embeddings_names, function(nm) {
      print(nm)
      
      # only use cells with labels
      data_all = get(paste0(nm, "_corrected"))
      labels_train = referenceLabels[!is.na(referenceLabels)]
      
      knn_out = embeddingKNN(data_all,
                             labels_train,
                             type = "uniform_fixed",
                             k_values = 5)
      
      acc = mean(isEqual(knn_out[names(queryLabels),"predicted_labels"], queryLabels), na.rm = TRUE)
      
      return(acc)
    }, simplify = FALSE)
    
    # calculate the resubstitution accuracy
    embeddings_resub_accuracy = sapply(embeddings_names, function(nm) {
      print(nm)
      
      # only use cells with labels
      data_all = get(paste0(nm, "_corrected"))
      labels_train = referenceLabels[!is.na(referenceLabels)]
      
      knn_out = embeddingKNN(data_all,
                             labels_train,
                             type = "uniform_fixed",
                             k_values = 5)
      
      acc = mean(isEqual(knn_out[names(labels_train),"resubstituted_labels"],
                         labels_train), na.rm = TRUE)
      return(acc)
    }, simplify = FALSE)
    
  
    res = rbind(res,
                data.frame(
                  genes = rep(length(genes), length(embeddings_names)),
                  type = names(embeddings_names),
                  Accuracy = unlist(embeddings_accuracy),
                  Accuracy_resub = unlist(embeddings_resub_accuracy),
                  querySample = rep(querySample, length(embeddings_names)),
                  Sim = sim
                ))
    
  }
}

dim(res)
```

Save the results

```{r}
saveRDS(res, file = "../../output/MGD_StabMap_example/MGD_StabMap_example_res.Rds")
saveRDS(simGenes, file = "../../output/MGD_StabMap_example/MGD_StabMap_example_simGenes.Rds")
```

Examine results

```{r}
  g = ggplot(res, aes(x = genes, y = Accuracy)) + 
      theme_classic() +
      geom_point(aes(colour = type)) + 
      geom_smooth(aes(group = type, colour = type), fill = NA, method = "loess") + 
      NULL
    print(g)
    
    g = ggplot(res, aes(x = genes, y = Accuracy_resub)) + 
      theme_classic() +
      geom_point(aes(colour = type)) + 
      geom_smooth(aes(group = type, colour = type), fill = NA, method = "loess") + 
      NULL
    print(g)
    
    g = ggplot(res, aes(x = factor(genes), y = Accuracy)) + 
      theme_classic() +
      stat_summary(aes(width = 0.8, fill = type), geom = "bar", fun = "mean", position = "dodge") +
      stat_summary(aes(group = type), geom = "errorbar", fun.data = "mean_se", position = "dodge", width = 0.8) +
      xlab("Number of genes") +
      labs(fill = "") +
      NULL
    print(g)
```

Finish

```{r}
sessionInfo()
```
