---
title: "StabMap: Mouse Gastrulation chimera and Spatial Mouse Atlas"
author: "Shila Ghazanfar"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
       html_document:
                     toc: true
                     toc_float:
                           collapsed: false
                           smooth_scroll: false
                     code_folding: hide
                     fig_width: 10 
                     fig_height: 8
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE,
                      cache = FALSE, cache.lazy = FALSE)
```

```{r}
set.seed(2021)
```

Load scripts and packages.

```{r}
source("../scripts/initialise.R")
library(gtools)
source("../../LohoffSeqFISH/celltype_colours.R")
```

Load scRNA-seq data SCE object.

```{r}
scrna = readRDS("../../MouseGastrulationData/MGD_chimera_scrna.Rds")
scrna
```

Perform feature selection on these.

```{r}
# Feature selection
decomp <- modelGeneVar(scrna, block = scrna$sample)
plot(decomp$mean, decomp$total)
hvgs <- rownames(decomp)[decomp$mean>0.01 & decomp$p.value <= 0.05]
points(decomp[hvgs,]$mean, decomp[hvgs,]$total, col = "blue")
length(hvgs)
scrna <- scrna[hvgs,]

rownames(scrna) <- make.unique(rowData(scrna)$SYMBOL)
```

Load seqFISH and smoothed expression data SCE object.

```{r}
seqFISH = readRDS("../../LohoffSeqFISH/seqFISH_smooth.Rds")
seqFISH
```

Integrate seqFISH and scRNA-seq only using StabMap and PCA.

```{r}
assay_list = list(
  seqFISH = as.matrix(logcounts(seqFISH)),
  scrna = as.matrix(logcounts(scrna))
)

mosaicDataUpSet(assay_list)
plot(mosaicDataTopology(assay_list))

meta <- colData(seqFISH)

SC_coord = setNames(c(meta[,"embryo"], rep("scrna", ncol(scrna))),
                    c(rownames(meta), colnames(scrna)))
table(SC_coord)

SC_coord_z = setNames(c(interaction(meta[,"embryo"], meta[,"z"]), rep("scrna", ncol(scrna))),
                    c(rownames(meta), colnames(scrna)))
table(SC_coord_z)

SC_sample = setNames(c(meta[,"embryo"], scrna$sample),
                    c(rownames(meta), colnames(scrna)))

SC_batch = setNames(c(as.character(meta$embryo_pos_z), scrna$sample),
                    c(rownames(meta), colnames(scrna)))
table(SC_batch)

SC_type = setNames(c(rep("seqFISH", nrow(meta)), rep("scrna", ncol(scrna))),
                   c(rownames(meta), colnames(scrna)))
table(SC_type)

cols = setNames(c(celltype_colours[as.character(meta[,"celltype_mapped_refined"])],
                  celltype_colours[as.character(scrna$celltype)]),
                c(rownames(meta), colnames(scrna))
)

celltypes = setNames(c(as.character(meta[,"celltype_mapped_refined"]),
                       as.character(scrna$celltype)),
                     c(rownames(meta), colnames(scrna))
)
```

Identify the batch correction ordering.

```{r}
batches_unsorted = unique(SC_batch)
batches_sorted = c(mixedsort(grep("atlas", batches_unsorted, value = TRUE)),
                   mixedsort(grep("wt_chim", batches_unsorted, value = TRUE)),
                   mixedsort(grep("^t_chim", batches_unsorted, value = TRUE)),
                   mixedsort(grep("embryo", batches_unsorted, value = TRUE))
                   )
batches_order = match(batches_sorted, batches_unsorted)
```

Perform naive PCA integration and batch correction using MNN.

```{r}
PC = mapPCA(assay_list)
colnames(PC) <- paste0("joint_PC", seq_len(ncol(PC)))
PC_corrected = reducedMNN_batchFactor(PC, SC_batch, merge.order = batches_order)
PC_corrected_umap = calculateUMAP_rnames(PC_corrected)

plot(PC_corrected_umap, asp = 1)
```

Perform StabMap and batch correction using MNN.

```{r}
SC_out = stabMap(assay_list,
                 reference_list = c("scrna", "seqFISH"),
                 projectAll = TRUE,
                 plot = FALSE,
                 scale.center = TRUE,
                 scale.scale = TRUE)

SC = reWeightEmbedding(SC_out, 
                       weights = list("seqFISH_PC" = 1,
                                      "scrna_PC" = 1
                       ))

SC_corrected = reducedMNN_batchFactor(SC, SC_batch, merge.order = batches_order)
SC_corrected_umap = calculateUMAP_rnames(SC_corrected)

plot(SC_corrected_umap, asp = 1)
```

```{r}
# coord_x = setNames(c(meta[,"x_global_affine"],
#                      reducedDim(scrna, "umap")[,1]),
#                    c(rownames(meta), colnames(scrna))
#                    )
# 
# coord_y = setNames(c(-meta[,"y_global_affine"],
#                      reducedDim(scrna, "umap")[,2]),
#                    c(rownames(meta), colnames(scrna))
#                    )

coord_x = setNames(c(meta[,"x_global_affine"],
                     rep(NA, ncol(scrna))),
                   c(rownames(meta), colnames(scrna))
                   )

coord_y = setNames(c(-meta[,"y_global_affine"],
                       rep(NA, ncol(scrna))),
                   c(rownames(meta), colnames(scrna))
                   )
```

Generate output files.

```{r}
SC_corrected_list = list(
  StabMap_uncorrected = SC,
  StabMap_corrected = SC_corrected,
  StabMap_umap = SC_corrected_umap,
  StabMap_meta = data.frame(cell = rownames(SC_corrected_umap),
                            sample = SC_coord[rownames(SC_corrected_umap)],
                            sample_z = SC_coord_z[rownames(SC_corrected_umap)],
                            batch = SC_batch[rownames(SC_corrected_umap)],
                            type = SC_type[rownames(SC_corrected_umap)],
                            celltype = celltypes[rownames(SC_corrected_umap)],
                            colour = cols[rownames(SC_corrected_umap)],
                            coord_x = coord_x[rownames(SC_corrected_umap)],
                            coord_y = coord_y[rownames(SC_corrected_umap)]),
  PC_uncorrected = PC,
  PC_corrected = PC_corrected,
  PC_umap = PC_corrected_umap,
  scrna_meta = colData(scrna),
  seqFISH_meta = colData(seqFISH)
)
saveRDS(SC_corrected_list, file = "../../output/MGA_chimera_seqFISH_StabMap_list.Rds")
# saveRDS(assay_list, file = "../../output/MGA_chimera_seqFISH_StabMap_assay_list.Rds")
```

# Finish

```{r}
sessionInfo()
```
