---
title: "StabMap simulation: PBMC disjoint mosaic integration with Multihop data"
author: "Shila Ghazanfar"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
       html_document:
                     toc: true
                     toc_float:
                           collapsed: false
                           smooth_scroll: false
                     code_folding: hide
                     fig_width: 10 
                     fig_height: 8
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE,
                      cache = FALSE, cache.lazy = FALSE)
```

```{r}
set.seed(2021)
```

Load scripts and packages.

```{r}
library(nord) # for colours
library(ggpubr) # for extracting legends
library(lisi) # installed via devtools::install_github("immunogenomics/LISI")
library(scattermore)
source("../scripts/initialise.R")
```

```{r}
modality_colours = setNames(nord(palette = "algoma_forest", 5)[c(1,3,4)],
                            c("Multiome", "ECCITESeq", "CYTOF"))
```

Load data from previous generating scripts. These are in the form of 
SingleCellExperiment objects.

10X Multiome data. Rename CD7 to unmatch the transcript name from protein.

```{r}
Multiome = readRDS("../../PBMC_Multiome/PBMC_multiome_sce.Rds")
Multiome
rnames = rownames(Multiome)
rnames[rnames == "CD7"] <- "RNA_CD7"
rnames[rnames == "MS4A1"] <- "RNA_MS4A1"
rownames(Multiome) <- rnames
```

CITE-seq (ECCITE-Seq) data. Rename the first appearing names as RNA to avoid
non-unique rownames.

```{r}
ECCITESeq = readRDS("../../ECCITEseq/eccite_sce.Rds")
ECCITESeq
rnames = rownames(ECCITESeq)
rnames_dup = rnames[duplicated(rnames)]
for (rnm in rnames_dup) {
  rnames[rnames == rnm & !duplicated(rnames)] <- paste0("RNA_", rnm)
}
any(duplicated(rnames))
rownames(ECCITESeq) <- rnames
```

CYTOF data. Rename protein to match. Add column names. Select the reference
sample with the largest number of cells.

```{r}
CYTOF = readRDS("../../CYTOF/cytof_PBMC_sce.Rds")
CYTOF
colnames(CYTOF) <- paste0("CYTOF_cell",seq_len(ncol(CYTOF)))

protName = c("CD20" = "MS4A1",
             "HLA-DR" = "MHCII (HLA-DR)")
rnames = setNames(rownames(CYTOF), rownames(CYTOF))
rnames[names(protName)] <- protName
rownames(CYTOF) <- unname(rnames)

nCells_CYTOF = unclass(table(CYTOF$sample_id, CYTOF$population_id))
sampleToKeep = grep("Reference", names(sort(rowSums(nCells_CYTOF), decreasing = TRUE)), value = TRUE)[1]

CYTOF <- CYTOF[,CYTOF$sample_id == sampleToKeep]
```

Examine the mosaic data topology.

```{r}
assay_list = lapply(list(CYTOF = logcounts(CYTOF),
                         ECCITESeq = logcounts(ECCITESeq),
                         Multiome = logcounts(Multiome)),as.matrix)

# this should be empty
intersect(intersect(rownames(Multiome), rownames(ECCITESeq)), rownames(CYTOF))

intersect(rownames(ECCITESeq), rownames(CYTOF))
setdiff(rownames(ECCITESeq), rownames(CYTOF))
setdiff(rownames(CYTOF), rownames(ECCITESeq))

any(duplicated(unlist(lapply(assay_list, colnames))))
lapply(assay_list, function(x) any(duplicated(rownames(x))))
lapply(assay_list, function(x) head(colnames(x)))

modality = setNames(rep(names(assay_list), times = lapply(assay_list, ncol)),
                    unlist(lapply(assay_list, colnames)))
```

Perform StabMap, employing multi-hop from CYTOF dataset to ECCITE dataset 
finally to Multiome dataset, and vice versa, and finally using both together.

```{r}
out_CYTOF = stabMap(assay_list,
                    reference_list = c("CYTOF"),
                    projectAll = TRUE,
                    ncomponentsReference = 10,
                    restrictFeatures = TRUE,
                    plot = FALSE)

out_Multiome = stabMap(assay_list,
                    reference_list = c("Multiome"),
                    projectAll = TRUE,
                    plot = FALSE,
                    restrictFeatures = TRUE)

out_joint = stabMap(assay_list,
                    reference_list = c("CYTOF", "Multiome"),
                    ncomponentsReference = c(CYTOF = 10, Multiome = 50),
                    projectAll = TRUE,
                    plot = FALSE,
                    restrictFeatures = TRUE)

# out_joint = cbind(out_CYTOF, out_Multiome[rownames(out_CYTOF),])
out_joint = reWeightEmbedding(out_joint, weights = c("CYTOF_PC" = 1, "Multiome_PC" = 1))

out_joint_corrected = reducedMNN_batchFactor(
  as.matrix(out_joint), batchFactor = modality[rownames(out_joint)])
out_CYTOF_corrected = reducedMNN_batchFactor(as.matrix(out_CYTOF), batchFactor = modality[rownames(out_CYTOF)])
out_Multiome_corrected = reducedMNN_batchFactor(as.matrix(out_Multiome), batchFactor = modality[rownames(out_Multiome)])

out_joint_UMAP_corrected = calculateUMAP_rnames(out_joint_corrected)
out_CYTOF_UMAP_corrected = calculateUMAP_rnames(out_CYTOF_corrected)
out_Multiome_UMAP_corrected = calculateUMAP_rnames(out_Multiome_corrected)

out_joint_df = data.frame(cell = names(modality),
                          modality = modality,
                          UMAP1_CYTOF_corrected = out_CYTOF_UMAP_corrected[names(modality),1],
                          UMAP2_CYTOF_corrected = out_CYTOF_UMAP_corrected[names(modality),2],
                          
                          UMAP1_Multiome_corrected = out_Multiome_UMAP_corrected[names(modality),1],
                          UMAP2_Multiome_corrected = out_Multiome_UMAP_corrected[names(modality),2],
                          
                          UMAP1_joint_corrected = out_joint_UMAP_corrected[names(modality),1],
                          UMAP2_joint_corrected = out_joint_UMAP_corrected[names(modality),2],
                          PC1_uncorrected = out_joint[names(modality),1],
                          PC2_uncorrected = out_joint[names(modality),2])
ctype_joint = c(setNames(as.character(Multiome$celltype), colnames(Multiome)),
                setNames(as.character(CYTOF$population_id), colnames(CYTOF)),
                setNames(rep("Unknown", ncol(ECCITESeq)), colnames(ECCITESeq)))
out_joint_df$celltype <- ctype_joint[rownames(out_joint_df)]
out_joint_df$sample_id <- colData(CYTOF)[rownames(out_joint_df),"sample_id"]
out_joint_df <- permuteRows(out_joint_df)

broad_celltypes = c("CD4 T-cells" = "CD4 T", "CD8 T-cells" = "CD8 T", "CD56 (bright) NK cells" = "NK", "classical monocytes" = "Monocyte", "memory CD4 T cells" = "CD4 T", "intermediate monocytes" = "Monocyte", "monocytes" = "Monocyte", "NK cells" = "NK", "B-cells IgM-" = "B", "non-classical monocytes" = "Monocyte", "naive B cells" = "B", "CD56 (dim) NK cells" = "NK", "memory B cells" = "B", "Unknown" = "Unknown", "naive CD4 T cells" = "CD4 T", "naive CD8 T cells" = "CD8 T", "myeloid DC" = "DC", "plasmacytoid DC" = "DC", "DC" = "DC", "B-cells IgM+" = "B", "effector CD8 T cells" = "CD8 T", "MAIT T cells" = "MAIT T", "surface-" = "Surface -")
out_joint_df$broad_celltype = broad_celltypes[out_joint_df$celltype]
```


```{r}
# export some plots:
g1 = ggplot(out_joint_df, aes(x = modality)) + 
  # geom_bar(fill = "gray") +
  geom_bar(aes(fill = modality)) + 
  scale_fill_manual(values = modality_colours) +
  ylab("Number of cells") +
  theme_classic() +
  xlab("Modality") +
  theme(legend.position = "none") +
  theme(axis.text = element_text(size = 15)) +
  theme(axis.title = element_text(size = 20)) +
  NULL
g1
ggsave(g1, file = paste0("../../Figures/raw/PBMC_multihop_numCells.pdf"),
       width = 5, height = 4)

g2 = mosaicDataUpSet(assay_list,
                     text.scale = c(3,1,2,1,3,3), 
                     point.size = 2.8, 
                     line.size = 1,
                     sets.x.label = "Cells",
                     mainbar.y.label = "Features",
                     main.bar.color = method_colours["StabMap"],
                     set_size.numbers_size = TRUE,
                     keep.order = TRUE,
                     sets = names(assay_list))
g2
pdf(paste0("../../Figures/raw/PBMC_multihop_upset.pdf"), 
    width = 5, height = 5)
print(g2)
dev.off()
  
graph = mosaicDataTopology(assay_list)
V(graph)$size = 70
V(graph)$color = modality_colours[V(graph)$name]
V(graph)$label.color = c("white", "black", "black")
plot(graph, layout = layout.drl)
  
pdf(paste0("../../Figures/raw/PBMC_multihop_network.pdf"), 
    width = 6, height = 6)
plot(graph, layout = layout.drl)
dev.off()
```


```{r}
g3List = sapply(c("CYTOF", "Multiome", "joint"), function(nm){
  g3_joint = ggplot(out_joint_df, aes(x = get(paste0("UMAP1_", nm, "_corrected")), y = get(paste0("UMAP2_", nm, "_corrected")))) + 
  geom_scattermore(aes(colour = modality), pointsize = 1.3, alpha = 0.5) +
  scale_colour_manual(values = modality_colours) +
  theme_classic() + 
  theme(legend.position = "bottom") +
  theme(axis.ticks = element_blank()) +
  theme(axis.text = element_blank()) +
  guides(colour = guide_legend(override.aes = list(size = 5, alpha = 1))) +
    ggtitle(nm) +
  coord_fixed() +
  xlab("UMAP1") + 
  ylab("UMAP2") +
  labs(colour = "") + 
  NULL
}, simplify = FALSE)

g4List = sapply(c("CYTOF", "Multiome", "joint"), function(nm){
    ggplot(subset(out_joint_df, broad_celltype != "Unknown"), 
           aes(x = get(paste0("UMAP1_", nm, "_corrected")),
               y = get(paste0("UMAP2_", nm, "_corrected")))) + 
  geom_scattermore(aes(colour = broad_celltype), pointsize = 1.3, alpha = 0.5) +
  # scale_colour_nord(palette = "lumina") +
  # scale_colour_manual(values = colors) +
  scale_colour_brewer(palette = "Set1") +
  theme_classic() + 
  theme(legend.position = "right",) +
  theme(axis.ticks = element_blank()) +
  theme(axis.text = element_blank()) +
  guides(colour = guide_legend(override.aes = list(size = 5, alpha = 1),
                               ncol = 1)) +
  coord_fixed() +
  xlab("UMAP1") + 
  ylab("UMAP2") +
  labs(colour = "") +
  NULL
}, simplify = FALSE)

wrap_plots(c(g3List, g4List))  + plot_layout(nrow = 2, ncol = 3)
```

Export these plots

```{r}
g = g3List[["CYTOF"]] + 
  theme(legend.position = "none") +
  ggtitle("Reference dataset: CYTOF")
ggsave(g, file = "../../Figures/raw/PBMC_multihop_CYTOF_modality.pdf",
       height = 6, width = 6)
g = g3List[["joint"]] + 
  theme(legend.position = "none") +
  ggtitle("Reference datasets: CYTOF and Multiome")
ggsave(g, file = "../../Figures/raw/PBMC_multihop_both_modality.pdf",
       height = 6, width = 6)
g = g3List[["Multiome"]] + 
  theme(legend.position = "none") +
  ggtitle("Reference dataset: Multiome")
ggsave(g, file = "../../Figures/raw/PBMC_multihop_Multiome_modality.pdf",
       height = 6, width = 6)
g = g4List[["CYTOF"]] + 
  theme(legend.position = "none") +
  ggtitle("")
ggsave(g, file = "../../Figures/raw/PBMC_multihop_CYTOF_celltype.pdf",
       height = 6, width = 6)
g = g4List[["joint"]] + 
  theme(legend.position = "none") +
  ggtitle("")
ggsave(g, file = "../../Figures/raw/PBMC_multihop_both_celltype.pdf",
       height = 6, width = 6)
g = g4List[["Multiome"]] + 
  theme(legend.position = "none") +
  ggtitle("")
ggsave(g, file = "../../Figures/raw/PBMC_multihop_Multiome_celltype.pdf",
       height = 6, width = 6)

g_leg = as_ggplot(get_legend(g3List[["CYTOF"]] + theme(legend.position = "right")))
g_leg
ggsave(g_leg, 
       file = paste0("../../Figures/raw/PBMC_multihop_UMAP_modality_legend.pdf"),
       height = 1, width = 3)

g_leg = as_ggplot(get_legend(g4List[["CYTOF"]]))
g_leg
ggsave(g_leg, 
       file = paste0("../../Figures/raw/PBMC_multihop_UMAP_celltype_legend.pdf"),
       height = 3, width = 3)
```

# Report LISI as measure of mixing

```{r}
lisi_cells = rownames(subset(out_joint_df,
                             modality %in% c("CYTOF", "Multiome")))
perpval = 100

lisi_values = compute_lisi(out_CYTOF_corrected[lisi_cells,],
                           out_joint_df[lisi_cells,],
                           c("modality", "broad_celltype"),
                           perplexity = perpval)
colnames(lisi_values) <- c("modality_CYTOF", "celltype_CYTOF")

lisi_values_Multiome = compute_lisi(out_Multiome_corrected[lisi_cells,],
                                    out_joint_df[lisi_cells,],
                                    c("modality", "broad_celltype"),
                           perplexity = perpval)
lisi_values$modality_Multiome <- lisi_values_Multiome$modality
lisi_values$celltype_Multiome <- lisi_values_Multiome$broad_celltype

lisi_values_both = compute_lisi(out_joint_corrected[lisi_cells,],
                                out_joint_df[lisi_cells,],
                                c("modality", "broad_celltype"),
                           perplexity = perpval)
lisi_values$modality_both <- lisi_values_both$modality
lisi_values$celltype_both <- lisi_values_both$broad_celltype

lisi_values_long = tidyr::pivot_longer(lisi_values,
                                       cols = colnames(lisi_values))
lisi_values_long$type <- ifelse(grepl("modality", lisi_values_long$name),
                                "Modality","Celltype")
lisi_values_long$reference <- gsub("modality_|celltype_", "", lisi_values_long$name)
lisi_values_long$name <- factor(lisi_values_long$name, levels = c(outer(c("celltype_", "modality_"),
                                                                       c("CYTOF", "Multiome", "both"),
                                                                      paste0)))

g = ggplot(lisi_values_long, aes(x = name, y = value)) + 
  geom_violin(aes(fill = reference), draw_quantiles = c(0.5), scale = "width") + 
  facet_wrap(~type, scales = "free", ncol = 2) + 
  theme_classic() + 
  theme(strip.background = element_rect(colour = NA)) +
  theme(strip.text = element_text(size = 14)) +
  xlab("") + 
  ylab("LISI value") +
  theme(legend.position = "none") +
  scale_fill_manual(values = c(modality_colours[c("CYTOF", "Multiome")], "both" = "grey"), 
                    labels = c("CYTOF", "Multiome", "Both"), drop = TRUE) +
  guides(fill = guide_legend(title = "Reference dataset")) +
  scale_x_discrete(labels = function(x) gsub("modality_|celltype_", "", ifelse(grepl("both", x), "Both", x))) +
  NULL
g

ggsave(g, file = "../../Figures/raw/PBMC_multihop_LISI_violin.pdf",
       height = 3, width = 6)

g_leg = as_ggplot(get_legend(g + theme(legend.position = "bottom")))
g_leg
ggsave(g_leg, file = "../../Figures/raw/PBMC_multihop_LISI_violin_leg.pdf",
       height = 1, width = 6)

```

Finish

```{r}
sessionInfo()
```