---
title: "StabMap simulation: PBMC Multiome de-matched cells"
author: "Shila Ghazanfar"
date: "23/09/2021"
output:
       html_document:
                     toc: true
                     toc_float:
                           collapsed: false
                           smooth_scroll: false
                     code_folding: hide
                     fig_width: 10 
                     fig_height: 8
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE,
                      cache = TRUE, cache.lazy = FALSE)
```

```{r}
set.seed(2021)
```

Load scripts and packages.

```{r}
library(SingleCellMultiModal)
library(reshape)
library(nord)
library(ggpubr)
library(Matrix)
library(scattermore)
source("../scripts/initialise.R")
```

# ATAC normalisation function

Normalise ATAC by binarisation, followed by TF-IDF normalisation, function with 
thanks to Ricard Argelaguet:

```{r}
# mtx is the binary peak matrix with dim (npeaks,ncells)
#'  method=1: The TF-IDF implementation used by Stuart & Butler et al. 2019. This computes \eqn{\log(TF \times IDF)}.
#'  method=2: The TF-IDF implementation used by Cusanovich & Hill et al. 2018. This computes \eqn{TF \times (\log(IDF))}.
#'  method=3: The log-TF method used by Andrew Hill. This computes \eqn{\log(TF) \times \log(IDF)}.
tfidf <- function(mtx, method = 1, scale.factor = 1e4) {
  npeaks <- colSums(mtx)
  if (any(npeaks == 0)) {
    warning("Some cells contain 0 total counts")
  }
  tf <- tcrossprod(mtx, y = Diagonal(x=1/npeaks))
  rsums <- rowSums(mtx)
  if (any(rsums == 0)) {
    warning("Some features contain 0 total counts")
  }
  idf <- ncol(mtx) / rsums
  if (method == 2) {
    idf <- log(1 + idf)
  } else if (method == 3) {
    tf <- log1p(tf * scale.factor)
    idf <- log(1 + idf)
  }
  mtx.tfidf <- Diagonal(n = length(idf), x = idf) %*% tf
  if (method == 1) {
    mtx.tfidf <- log1p(mtx.tfidf * scale.factor)
  }
  colnames(mtx.tfidf) <- colnames(mtx)
  rownames(mtx.tfidf) <- rownames(mtx)
  # set NA values to 0
  mtx.tfidf[is.na(mtx.tfidf)] <- 0
  return(mtx.tfidf)
}
```

Data using Bioconductor SingleCellMultiModal package

```{r}
mae <- scMultiome("pbmc_10x", mode = "*", dry.run = FALSE, format = "MTX")

colors <- viridis::viridis(length(unique(mae$celltype)))
names(colors) <- sort(unique(as.character(mae$celltype)))

modality_colours = nord(palette = "algoma_forest", 5)[c(1,3,4)]
names(modality_colours) = c("RNA", "ATAC", "Multiome")

mae

upsetSamples(mae)

head(colData(mae))

dim(experiments(mae)[["rna"]])

names(experiments(mae))

sce.rna <- experiments(mae)[["rna"]]

# Normalisation
sce.rna <- logNormCounts(sce.rna)

# Feature selection
decomp <- modelGeneVar(sce.rna)
hvgs <- rownames(decomp)[decomp$mean>0.01 & decomp$p.value <= 0.05]

length(hvgs)

sce.rna <- sce.rna[hvgs,]

# PCA
sce.rna <- scater::runPCA(sce.rna, ncomponents = 25)

# UMAP
set.seed(42)
sce.rna <- scater::runUMAP(sce.rna, dimred="PCA", n_neighbors = 25, min_dist = 0.3)
plotUMAP(sce.rna, colour_by="celltype", point_size=0.5, point_alpha=1)
```

Peak annotation downloaded from the MOFA vignette, file available here ftp://ftp.ebi.ac.uk/pub/databases/mofa/10x_rna_atac_vignette/atac_peak_annotation.tsv
The MOFA+ vignette itself is available here
https://raw.githack.com/bioFAM/MOFA2_tutorials/master/R_tutorials/10x_scRNA_scATAC.html

Identify peaks located in promoter regions of known genes.

```{r}
peakInfo = read.delim("../../PBMC_Multiome/atac_peak_annotation.tsv", header = TRUE, row.names = 1)
peakNames = rownames(peakInfo)

ind = (gsub(";.*", "", peakInfo[, "gene"]) %in% rowData(sce.rna)[,"id"]) &
  (peakInfo$peak_type == "promoter")

peakInfo$gene_new <- rownames(peakInfo)
peakInfo[ind,"gene_new"] <- peakInfo[ind,"gene"]
```

```{r}
dim(experiments(mae)[["atac"]])

sce.atac <- experiments(mae)[["atac"]]

# Normalise using TF-IDF
logcounts(sce.atac) <- tfidf(1*(counts(sce.atac)>0))

# Feature selection using highly variable peaks
# And adding matching peaks to genes
decomp <- modelGeneVar(sce.atac)
plot(decomp$mean, decomp$total)
hvgs <- rownames(decomp)[decomp$mean>0.25
                         & decomp$p.value <= 0.05
                         ]
length(hvgs)
# ATAC HVGs and all promoter peaks of RNA HVGs
hvgs_all = union(hvgs, intersect(rownames(peakInfo)[ind],
                                 rownames(decomp)[decomp$mean>0.25
                         ]))
length(hvgs_all)
sce.atac <- sce.atac[hvgs_all,]

# PCA of ATAC modality
sce.atac <- scater::runPCA(sce.atac, ncomponents = 50)

# UMAP
set.seed(42)
sce.atac <- scater::runUMAP(sce.atac, dimred="PCA", n_neighbors = 25, min_dist = 0.3)
plotUMAP(sce.atac, colour_by="celltype", point_size=0.5, point_alpha=1)
```

Match feature names for mosaic data integration

```{r}
length(intersect(rownames(sce.atac), rownames(sce.rna)))

rnames.atac = peakInfo[rownames(sce.atac), "gene_new"]
rnames.rna = rowData(sce.rna)[,"id"]

length(intersect(rnames.atac, rnames.rna))
length(setdiff(rnames.atac, rnames.rna))
length(setdiff(rnames.rna, rnames.atac))

rownames(sce.rna) <- rnames.rna
rownames(sce.atac) <- rnames.atac

length(intersect(rownames(sce.atac), rownames(sce.rna)))

nrow(sce.atac)
length(unique(rownames(sce.atac)))
sce.atac <- sce.atac[!duplicated(rownames(sce.atac)),]

nrow(sce.rna)
length(unique(rownames(sce.rna)))
sce.rna <- sce.rna[!duplicated(rownames(sce.rna)),]

# give the ATAC cells a new suffix
colnames(sce.atac) <- paste0(colnames(sce.atac), "_atac")

length(intersect(rownames(sce.atac), rownames(sce.rna)))
```

Build input data for StabMap and other methods.

```{r}
SCE_list = list(atac = sce.atac,
                rna = sce.rna)
assayNames = list(atac = "logcounts",
                  rna = "logcounts")
assay_list = mapply(assay, SCE_list, assayNames)
counts_list = mapply(assay, SCE_list, "counts")
labels_list = list(rna = sce.rna$celltype)
grouping = "celltype"
k = 50

ct_all = unlist(lapply(lapply(SCE_list, colData), "[", , grouping))
names(ct_all) <- unlist(lapply(lapply(SCE_list, colData), rownames))

celltype_colours = colors

plotFeatureOverlaps(assay_list)

g = plotFeatureOverlaps(assay_list)
pdf(paste0("../../Figures/raw/PBMC_multiome_upset.pdf"), 
    width = 4, height = 3)
print(g)
dev.off()
```

Perform StabMap and other methods.

```{r}
# PCA embedding (naive)
PC_embedding = mapPCA(assay_list = assay_list,
                      nPCs = k)
dim(PC_embedding)

# UINMF embedding
UINMF_embedding = UINMF_wrap(counts_list = counts_list,
                             ncomponentsSubset = k)

# StabMap
SC_embedding = stabMapGeneralised(assay_list,
                                  projectAll = TRUE,
                                  scale.center = TRUE,
                                  scale.scale = TRUE)
dim(SC_embedding)

# Give equal weighting to RNA and ATAC
SC_embedding <- reWeightEmbedding(SC_embedding)

# MultiMAP
MultiMAP_embedding = MultiMAP_wrap(assay_list, verbose = TRUE)

embeddings_names = c("StabMap" = "SC_embedding",
                     "PCA" = "PC_embedding",
                     "UINMF" = "UINMF_embedding",
                     "MultiMAP" = "MultiMAP_embedding"
)

# these methods already perform batch correction
no_batch = c("MultiMAP_embedding", "UINMF_embedding")
# this method already gives 2D embedding
no_umap = c("MultiMAP_embedding")
```

Batch correct and plot.

```{r, include=FALSE}
type = factor(ifelse(grepl("atac", rownames(PC_embedding)),
                     "ATAC", "RNA"))
names(type) <- rownames(PC_embedding)
table(type)
batchFactor = type

sapply(embeddings_names, function(nm) {
  print(nm)
  if (nm %in% no_batch) {
    assign(paste0(nm, "_corrected"), get(nm), envir = .GlobalEnv)
  } else {
    assign(paste0(nm, "_corrected"), reducedMNN_batchFactor(get(nm), batchFactor),
           envir = .GlobalEnv)
  }
}, simplify = FALSE)
```

Plotting functions

```{r}
pl = function(embedding_UMAP, name, coltype = "type", xlab = "UMAP1", ylab = "UMAP2", title = TRUE, facet = FALSE, fixed = TRUE) {
  ind = sample(rownames(embedding_UMAP))
  df = data.frame(cell = ind,
                  UMAP1 = embedding_UMAP[ind,1],
                  UMAP2 = embedding_UMAP[ind,2],
                  type = factor(type[ind]),
                  ctype = ct_all[ind])
  df$col = df[,coltype]
  p = ggplot(df, aes(x = UMAP1, y = UMAP2, colour = col)) + 
    geom_scattermore(pointsize = 1.3, alpha = 0.5) +
    theme_classic() + 
    theme(axis.ticks = element_blank()) +
    theme(axis.text = element_blank()) +
    theme(legend.position = "none") +
    xlab(xlab) +
    ylab(ylab) +
    NULL
  
  if (fixed) {
    p <- p + coord_fixed()
  }
  
  if (title) {
    p <- p +  ggtitle(name)

  }
  
  if (coltype == "ctype") {
    p <- p + scale_colour_manual(values = colors)
  }
  
  if (coltype == "type") {
    p <- p + scale_colour_manual(values = modality_colours)
  }
  
   if (facet) {
          p <- p + facet_wrap(~type, nrow = 2)  +
            NULL
        }
  
  return(p)
}

# calculate UMAPs
embeddings_corrected_UMAP = sapply(embeddings_names, function(nm){
  if (nm %in% no_umap) return(get(nm)) # i.e. do nothing
  calculateUMAP_rnames(get(paste0(nm, "_corrected")))
}, simplify = FALSE)

p_corrected = 
  wrap_plots(
    c(
      mapply(pl, embeddings_corrected_UMAP, names(embeddings_corrected_UMAP), 
             ylab = c("UMAP2", rep("", length(embeddings_corrected_UMAP)-1)),
             MoreArgs = list(xlab = ""), SIMPLIFY = FALSE),
      mapply(pl, embeddings_corrected_UMAP, names(embeddings_corrected_UMAP), 
             ylab = c("UMAP2", rep("", length(embeddings_corrected_UMAP)-1)),
             MoreArgs = list(coltype = "ctype", title = FALSE), SIMPLIFY = FALSE)
    )) +
  plot_layout(nrow = 2, ncol = length(embeddings_corrected_UMAP), byrow = TRUE)
p_corrected
ggsave(p_corrected, file = "../../Figures/raw/PBMC_multiome_UMAPs.pdf",
       height = 4, width = 8)

g = pl(embeddings_corrected_UMAP[[1]], "") + 
  theme(legend.position = "bottom") +
  guides(colour = guide_legend(override.aes = list(size = 5, alpha = 1))) + 
  labs(colour = "")
  
g_leg = as_ggplot(get_legend(g))
g_leg
ggsave(g_leg, file = "../../Figures/raw/PBMC_multiome_UMAPs_modalitylegend.pdf",
       height = 2, width = 2)

g = pl(embeddings_corrected_UMAP[[1]], "", coltype = "ctype") + 
  theme(legend.position = "right") +
  guides(colour = guide_legend(override.aes = list(size = 5, alpha = 1))) + 
  labs(colour = "")
  
g_leg = as_ggplot(get_legend(g))
g_leg
ggsave(g_leg, file = "../../Figures/raw/PBMC_multiome_UMAPs_celltypelegend.pdf",
       height = 6, width = 2)

# also split cell type coloured plot by type
p_corrected = 
  wrap_plots(
    c(
      mapply(pl, embeddings_corrected_UMAP, names(embeddings_corrected_UMAP), 
             ylab = c("UMAP2", rep("", length(embeddings_corrected_UMAP)-1)),
             MoreArgs = list(coltype = "ctype", title = TRUE, facet = TRUE, 
                             fixed = FALSE), SIMPLIFY = FALSE)
    )) +
  plot_layout(nrow = 2, ncol = length(embeddings_corrected_UMAP), byrow = TRUE)
p_corrected

ggsave(p_corrected, file = "../../Figures/raw/PBMC_multiome_UMAPs_split.pdf",
       height = 4, width = 8)
```

Predict cell type labels of the ATAC-derived cells.

```{r}
# treat RNA as the reference
referenceLabels = colData(sce.rna)[,grouping]
names(referenceLabels) = colnames(sce.rna)

queryLabels = colData(sce.atac)[,grouping]
names(queryLabels) = colnames(sce.atac)

# calculate accuracy of the query cells
embeddings_accuracy = sapply(embeddings_names, function(nm) {
  print(nm)
  
  # only use cells with labels
  data_all = get(paste0(nm, "_corrected"))
  labels_train = referenceLabels[!is.na(referenceLabels)]
  
  knn_out = embeddingKNN(data_all,
                         labels_train,
                         type = "uniform_fixed",
                         k_values = 5)
  
  acc = mean(isEqual(knn_out[names(queryLabels),"predicted_labels"], queryLabels), na.rm = TRUE)
  
  acc_bal = mean(unlist(lapply(split(isEqual(knn_out[names(queryLabels),"predicted_labels"], queryLabels), queryLabels), mean, na.rm = TRUE)))
  
  return(list(acc = acc, acc_bal = acc_bal))
}, simplify = FALSE)

res = data.frame(
  type = names(embeddings_names),
  Accuracy = unlist(lapply(embeddings_accuracy, "[[", "acc")),
  Accuracy_balanced = unlist(lapply(embeddings_accuracy, "[[", "acc_bal"))
)
res <- res[order(res$Accuracy, decreasing = TRUE),]
res$type <- factor(res$type, levels = res$type)

typesToKeep = c("MultiMAP",
                "PCA",
                "StabMap",
                "UINMF")

g1 = ggplot(subset(res, type %in% typesToKeep), aes(x = type, y = Accuracy_balanced)) + 
  theme_classic() +
  stat_summary(aes(width = 0.8, fill = type), geom = "bar", fun = "mean", position = "dodge") +
  labs(fill = "") +
  theme_classic() +
  NULL
g1
g2 = ggplot(subset(res, type %in% typesToKeep), aes(x = type, y = Accuracy)) + 
  theme_classic() +
  stat_summary(aes(width = 0.8, fill = type), geom = "bar", fun = "mean", position = "dodge") +
  labs(fill = "") +
  xlab("") +
  theme_classic() +
  ylab("Cell type accuracy") +
  scale_fill_manual(values = method_colours) +
  theme(legend.position = "none") +
  NULL
g2
ggsave(g2, file = "../../Figures/raw/PBMC_multiome_accuracy.pdf",
       height = 4, width = 4)
```

Calculate Jaccard indices of local neighbourhoods of each cell and modality.
Also compare with performing PCA separately per modality.

```{r}
embList = list()

for (k1 in c(30, 50, 100, 200)) {
  k2 = k1
  
  print(k1)
  
  embeddings_jaccard = sapply(embeddings_names, function(nm) {
    
    print(nm)
    
    embedding = get(paste0(nm, "_corrected"))
    # split into two matching matrices
    is_RNA = !grepl("atac", rownames(embedding))
    embedding_RNA = embedding[is_RNA,]
    embedding_ATAC = embedding[!is_RNA,]
    rownames(embedding_ATAC) <- gsub("_atac", "", rownames(embedding_ATAC))
    cells = intersect(rownames(embedding_RNA), rownames(embedding_ATAC))
    embedding_RNA_common = embedding_RNA[cells,]
    embedding_ATAC_common = embedding_ATAC[cells,]
    
    embedding_jaccard = embeddingJaccard(embedding_RNA_common, embedding_ATAC_common,
                                         k1 = k1, k2 = k2)
    return(embedding_jaccard)
  }, simplify = FALSE)
  
  # also want to compare against PC separate, using all data available
  pc_atac = calculatePCA(sce.atac)
  pc_rna = calculatePCA(sce.rna)
  rownames(pc_atac) <- gsub("_atac", "", rownames(pc_atac))
  cells = intersect(rownames(pc_atac), rownames(pc_rna))
  pc_rna_common = pc_rna[cells,]
  pc_atac_common = pc_atac[cells,]
  
  embeddings_jaccard[["PC_separate"]] <- embeddingJaccard(
    pc_rna_common, pc_atac_common,
    k1 = k1, k2 = k2)
  
  embList[[as.character(k1)]] <- embeddings_jaccard
}

df_jaccard_all = NULL

for (i in names(embList)) {
  
  embeddings_jaccard <- embList[[i]]  
  
  df_jaccard = data.frame(type = rep(names(embeddings_jaccard), times = unlist(lapply(embeddings_jaccard, length))),
                          jaccard = unlist(embeddings_jaccard),
                          k = i
  )
  
  df_jaccard_all <- rbind(df_jaccard_all, df_jaccard)
  
}

embeddings_jaccard <- embList[["50"]]

emb = sort(unlist(lapply(embeddings_jaccard, mean)))
emb

df_jaccard_all$k <- factor(df_jaccard_all$k, levels = gtools::mixedsort(unique(as.character(df_jaccard_all$k))))

df_jaccard_all$type <- factor(df_jaccard_all$type, levels = names(emb))

g = ggplot(df_jaccard_all, aes(x = k, y = jaccard, fill = type)) + 
  theme_classic() +
  geom_violin(draw_quantiles = c(0.5), adjust = 1.5, scale = "width") +
  labs(fill = "") +
  scale_fill_manual(values = method_colours) +
  ylab("Jaccard similarity") +
  theme(legend.position = "none") +
  xlab("") +
  NULL
print(g)
ggsave(g, file = "../../Figures/raw/PBMC_multiome_jaccard_all.pdf",
       height = 4, width = 6)

df_jaccard = data.frame(type = rep(names(embeddings_jaccard), times = unlist(lapply(embeddings_jaccard, length))),
                        jaccard = unlist(embeddings_jaccard)
)
df_jaccard$type <- factor(df_jaccard$type, levels = names(emb))

g = ggplot(subset(df_jaccard, type %in% c(typesToKeep, "PC_separate")), aes(x = type, y = jaccard, fill = type)) + 
  theme_classic() +
  geom_violin(draw_quantiles = c(0.5), adjust = 1.5, scale = "width") +
  labs(fill = "") +
  scale_fill_manual(values = method_colours) +
  ylab("Jaccard similarity among 50 neighbours") +
  theme(legend.position = "none") +
  xlab("") +
  NULL
print(g)
ggsave(g, file = "../../Figures/raw/PBMC_multiome_jaccard.pdf",
       height = 4, width = 6)
```

Calculate closest neighbour metric, i.e. how many other cells are between a 
cell and its matching cell with the other modality? Censor values at 501, i.e.
do not specifically calculate how many closest for any above that level.

```{r}
k_max = 501

nei_pos_list = sapply(embeddings_names, function(nm) {
  
  print(nm)
  
  embedding = get(paste0(nm, "_corrected"))
  
  nei_atac = queryNamedKNN(embedding,
                           embedding[grepl("_atac", rownames(embedding)),],
                           k = k_max)
  nei_atac_bin = apply(nei_atac, 2, function(x) isEqual(gsub("_atac", "", rownames(nei_atac)), x))
  nei_atac_pos = max.col(nei_atac_bin, ties.method = "last")
  mean(nei_atac_pos)
  nei_rna = queryNamedKNN(embedding,
                          embedding[!grepl("_atac", rownames(embedding)),],
                          k = k_max)
  nei_rna_bin = apply(nei_rna, 2, function(x) isEqual(paste0(rownames(nei_rna), "_atac"), x))
  nei_rna_pos = max.col(nei_rna_bin, ties.method = "last")
  mean(nei_rna_pos)
  
  nei_pos = c(nei_atac_pos, nei_rna_pos)
  names(nei_pos) <- c(rownames(nei_atac), rownames(nei_rna))
  
  return(nei_pos)
  
}, simplify = FALSE)

xvals = c(5,10,20,30,50,100,200,500)
nei_x = sapply(xvals,
               function(y) unlist(lapply(nei_pos_list, function(x) sum(x <= y))))
colnames(nei_x) <- xvals

nei_df = reshape::melt(nei_x)
nei_df$X1 <- factor(rownames(nei_x), levels = names(sort(nei_x[,1], decreasing = TRUE)))

g = ggplot(nei_df, aes(x = X2, y = value, group = X1)) + 
  theme_classic() +
  geom_line(aes(colour = X1), size = 1, alpha = 0.7) + 
  scale_y_log10() +
  scale_x_log10() +
  labs(colour = "", linetype = "") +
  scale_colour_manual(values = method_colours) +
  xlab("X") +
  ylab("Number of cells with at most X other\ncells nearer to matched cell") +
  theme(legend.position = "bottom") +
  NULL
g

ggsave(g, file = "../../Figures/raw/PBMC_multiome_matchingcell.pdf",
       height = 4, width = 4.5)

g = ggplot(nei_df, aes(x = factor(X2), y = value, group = X1)) + 
  theme_classic() +
  geom_col(aes(fill = X1, group = X1), size = 1, alpha = 0.7, position = "dodge") + 
  scale_y_log10() +
  geom_line(aes(colour = X1), size = 1, alpha = 0.7) + 
  labs(colour = "", linetype = "", fill = "") +
  scale_colour_manual(values = method_colours, aesthetics = c("colour", "fill")) +
  xlab("X") +
  ylab("Number of cells with at most X other\ncells nearer to matched cell") +
  theme(legend.position = "bottom") +
  NULL
g
ggsave(g + theme(legend.position = "none"), file = "../../Figures/raw/PBMC_multiome_matchingcell_bars.pdf",
       height = 4, width = 4.5)

g_leg = as_ggplot(get_legend(g))
g_leg
ggsave(g_leg, file = "../../Figures/raw/PBMC_multiome_matchingcell_bars_legend.pdf",
       height = 2, width = 6)
```

Aside: Split UMAP plot

```{r}

splitMultiomeUMAP = function(UMAP){
  is_RNA = !grepl("atac", rownames(UMAP))
  UMAP_RNA = UMAP[is_RNA,]
  UMAP_ATAC = UMAP[!is_RNA,]
  rownames(UMAP_ATAC) <- gsub("_atac", "", rownames(UMAP_ATAC))

  cells = intersect(rownames(UMAP_RNA), rownames(UMAP_ATAC))
  UMAP_split = cbind(UMAP_RNA[cells,], UMAP_ATAC[cells,])
  colnames(UMAP_split) <- c("UMAP1_RNA", "UMAP2_RNA", "UMAP1_ATAC", "UMAP2_ATAC")
  return(UMAP_split)
}


```

```{r}
library(ggiraph)
i = "StabMap"

df = as.data.frame(splitMultiomeUMAP(embeddings_corrected_UMAP[[i]]))
df$cell = rownames(df)
df$celltype = colData(sce.rna[,rownames(df)])[,grouping]

g0 = pl(embeddings_corrected_UMAP[[i]], i)

gg0 <- ggplot(df) +
  theme_classic() +
  geom_point_interactive(aes(x = UMAP1_RNA, y = UMAP2_RNA, 
                             tooltip = cell, data_id = cell), 
                         color = "red", size = .5) +
  geom_point_interactive(aes(x = UMAP1_ATAC, y = UMAP2_ATAC, 
                             tooltip = cell, data_id = cell), 
                         color = "blue", size = .5) +
  theme(legend.position = "none") + 
  ggtitle(paste0(i, " - all"))

gg1 <- ggplot(df) +
  theme_classic() +
  geom_point_interactive(aes(x = UMAP1_ATAC, y = UMAP2_ATAC), colour = "grey", size = .5) +
  geom_point_interactive(aes(x = UMAP1_RNA, y = UMAP2_RNA, color = celltype, 
                             tooltip = cell, data_id = cell), size = .5) +
  theme(legend.position = "none") + 
  ggtitle(paste0(i, " - RNA"))

gg2 <- ggplot(df) +
  theme_classic() +
  geom_point_interactive(aes(x = UMAP1_RNA, y = UMAP2_RNA), colour = "grey", size = .5) +
  geom_point_interactive(aes(x = UMAP1_ATAC, y = UMAP2_ATAC, color = celltype, 
                             tooltip = cell, data_id = cell), size = .5) +
  theme(legend.position = "none") +
  ggtitle(paste0(i, " - ATAC"))

g0 + gg1 + gg2

if (FALSE) { # can only be run interactively
  girafe(ggobj = plot_grid(gg0, gg1, gg2), width_svg = 8, height_svg = 4,
         options = list(
           opts_selection(type = "multiple",
                          css = "fill:#FF3333;stroke:black;",
                          only_shiny = FALSE)
         ))
}
```

# Finish

```{r}
sessionInfo()
```