---
title: "StabMap simulation: PBMC Multiome disjoint mosaic integration"
author: "Shila Ghazanfar"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
       html_document:
                     toc: true
                     toc_float:
                           collapsed: false
                           smooth_scroll: false
                     code_folding: hide
                     fig_width: 10 
                     fig_height: 8
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE,
                      cache = FALSE, cache.lazy = FALSE)
```

```{r}
set.seed(2022)
```

Load scripts and packages.

```{r}
library(SingleCellMultiModal)
library(MultiAssayExperiment)
library(nord) # for colours
library(ggpubr) # for extracting legends
library(lisi) # installed via devtools::install_github("immunogenomics/LISI")
library(scattermore)
source("../scripts/initialise.R")
```

Load data using SingleCellMultiModal Bioconductor package.

```{r}
mae <- scMultiome("pbmc_10x", mode = "*", dry.run = FALSE, format = "MTX")
```

Perform some exploration of this data.

```{r}
colors <- viridis::viridis(length(unique(mae$celltype)))
names(colors) <- sort(unique(as.character(mae$celltype)))

modality_colours = nord(palette = "algoma_forest", 5)[c(1,3,4)]
names(modality_colours) = c("RNA", "ATAC", "Multiome")

mae

upsetSamples(mae)

head(colData(mae))

dim(experiments(mae)[["rna"]])

names(experiments(mae))

sce.rna <- experiments(mae)[["rna"]]
```

Normalise and select features for RNA modality. Perform PCA and visualise the RNA modality.

```{r}
# Normalisation
sce.rna <- computeSumFactors(sce.rna)
sce.rna <- logNormCounts(sce.rna)

# Feature selection
decomp <- modelGeneVar(sce.rna)
hvgs <- rownames(decomp)[decomp$mean>0.01 & decomp$p.value <= 0.05]

length(hvgs)

sce.rna <- sce.rna[hvgs,]

# PCA
sce.rna <- scater::runPCA(sce.rna, ncomponents = 25)

# UMAP
sce.rna <- scater::runUMAP(sce.rna, dimred="PCA", n_neighbors = 25, min_dist = 0.3)
plotUMAP(sce.rna, colour_by="celltype", point_size=0.5, point_alpha=1)
```

Normalise and select features for ATAC modality. Perform PCA and visualise the ATAC modality. Note these cells are the same as the RNA modality.

```{r}
dim(experiments(mae)[["atac"]])

sce.atac <- experiments(mae)[["atac"]]

# Normalise using TF-IDF
logcounts(sce.atac) <- tfidf(1*(counts(sce.atac)>0))

# Feature selection using highly variable peaks
decomp <- modelGeneVar(sce.atac)
hvgs <- rownames(decomp)[decomp$mean>0.25 & decomp$p.value <= 0.05]
sce.atac <- sce.atac[hvgs,]

# PCA of ATAC modality
sce.atac <- scater::runPCA(sce.atac, ncomponents = 50)

# UMAP
sce.atac <- scater::runUMAP(sce.atac, dimred="PCA", n_neighbors = 25, min_dist = 0.3)
plotUMAP(sce.atac, colour_by="celltype", point_size=0.5, point_alpha=1)
```

Generate an assay_list

```{r}
probval = 1

modality = sample(c("RNA", "Multiome", "ATAC"), size = ncol(sce.rna),
                  replace = TRUE,
                  prob = c(probval,1,probval))
names(modality) <- colnames(sce.rna)
table(modality)

# multiome RNA
M_R = assay(sce.rna, "logcounts")[,modality == "Multiome"]
M_R_c = assay(sce.rna, "counts")[,modality == "Multiome"]

# multiome ATAC
M_A = assay(sce.atac, "logcounts")[,modality == "Multiome"]
M_A_c = assay(sce.atac, "counts")[,modality == "Multiome"]

# RNA
R = assay(sce.rna, "logcounts")[,modality == "RNA"]
R_c = assay(sce.rna, "counts")[,modality == "RNA"]

# ATAC
A = assay(sce.atac, "logcounts")[,modality == "ATAC"]
A_c = assay(sce.atac, "counts")[,modality == "ATAC"]

# they dont have any overlapping features
Reduce(intersect, list(c(rownames(M_R), rownames(M_A)), rownames(R), rownames(A)))

# there should be no overlapping colnames between the data,
# but colnames should be identical between the two multiome sub matrices
identical(colnames(M_R), colnames(M_A))
intersect(colnames(M_R), colnames(R))
intersect(colnames(M_R), colnames(A))
intersect(colnames(R), colnames(A))

assay_list_counts = list(RNA = R_c,
                         ATAC = A_c,
                         Multiome = rbind(M_R_c, M_A_c))

# logcounts
assay_list = list(RNA = R,
                  ATAC = A,
                  Multiome = rbind(M_R, M_A))
```


Split the cells into three groups with ratios provided below, first for RNA modality, second for 10X Multiome modality, and third for ATAC modality. Using this simulation, perform StabMap. Note that other techniques such as PCA, UINMF, or MultiMAP require some further matching of features between the RNA and ATAC modalities to be able to run.

Assess the quality of the mapping visually.

```{r}
lisi_values_list = list()
pred = NULL

for (probval in rev(c(1,2,3,4,5,10,20))) {
  
  print(probval)
  
  outFile = paste0("../../output/PBMC_nonoverlapping_multiome_", probval,
                   "_embeddings.Rds")
  if (file.exists(outFile)) {
    print("file exists, loading..")
    out = readRDS(outFile)
    list2env(out, envir = .GlobalEnv)
    # next
  } else {
    
    modality = sample(c("RNA", "Multiome", "ATAC"), size = ncol(sce.rna),
                      replace = TRUE,
                      prob = c(probval,1,probval))
    names(modality) <- colnames(sce.rna)
    table(modality)
    
    # multiome RNA
    M_R = assay(sce.rna, "logcounts")[,modality == "Multiome"]
    M_R_c = assay(sce.rna, "counts")[,modality == "Multiome"]
    
    # multiome ATAC
    M_A = assay(sce.atac, "logcounts")[,modality == "Multiome"]
    M_A_c = assay(sce.atac, "counts")[,modality == "Multiome"]
    
    # RNA
    R = assay(sce.rna, "logcounts")[,modality == "RNA"]
    R_c = assay(sce.rna, "counts")[,modality == "RNA"]
    
    # ATAC
    A = assay(sce.atac, "logcounts")[,modality == "ATAC"]
    A_c = assay(sce.atac, "counts")[,modality == "ATAC"]
    
    # they dont have any overlapping features
    Reduce(intersect, list(c(rownames(M_R), rownames(M_A)), rownames(R), rownames(A)))
    
    # there should be no overlapping colnames between the data,
    # but colnames should be identical between the two multiome sub matrices
    identical(colnames(M_R), colnames(M_A))
    intersect(colnames(M_R), colnames(R))
    intersect(colnames(M_R), colnames(A))
    intersect(colnames(R), colnames(A))
    
    assay_list_counts = list(RNA = R_c,
                             ATAC = A_c,
                             Multiome = rbind(M_R_c, M_A_c))
    
    # logcounts
    assay_list = list(RNA = R,
                      ATAC = A,
                      Multiome = rbind(M_R, M_A))
    
    out_joint = stabMap(assay_list,
                        projectAll = TRUE,
                        restrictFeatures = TRUE,
                        plot = FALSE,
                        scale.center = TRUE,
                        scale.scale = TRUE)
    out_joint <- reWeightEmbedding(out_joint, weights = c("RNA_PC" = 1, "ATAC_PC" = 1, "Multiome_PC" = 1))
    
    out_joint_corrected = reducedMNN_batchFactor(
      as.matrix(out_joint), batchFactor = modality[rownames(out_joint)])
    out_joint_UMAP_corrected = calculateUMAP_rnames(out_joint_corrected)
    
    out_joint_df = data.frame(cell = names(modality),
                              modality = modality,
                              celltype = colData(sce.rna)[names(modality),"celltype"],
                              UMAP1_corrected = out_joint_UMAP_corrected[names(modality),1],
                              UMAP2_corrected = out_joint_UMAP_corrected[names(modality),2])
    out_joint_df <- out_joint_df[sample(rownames(out_joint_df)),]
    
    # project to RNA space only - this is "multihop" mosaic integration
    out_joint_RNA = stabMap(assay_list,
                            reference_list = "RNA",
                            projectAll = TRUE,
                            restrictFeatures = TRUE,
                            plot = FALSE,
                            scale.center = TRUE,
                            scale.scale = TRUE)
    
    out_joint_RNA_corrected = reducedMNN_batchFactor(
      as.matrix(out_joint_RNA), batchFactor = modality[rownames(out_joint_RNA)])
    out_joint_RNA_UMAP_corrected = calculateUMAP_rnames(out_joint_RNA_corrected)
    
    out_joint_df$UMAP1_RNA <- out_joint_RNA_UMAP_corrected[rownames(out_joint_df),1]
    out_joint_df$UMAP2_RNA <- out_joint_RNA_UMAP_corrected[rownames(out_joint_df),2]
    
    # perform cobolt integration and add their UMAP coordinates
    # note that Cobolt wants count data not logcounts
    out_cobolt = cobolt_wrap(assay_list_counts, verbose = TRUE)
    out_cobolt_UMAP = calculateUMAP_rnames(out_cobolt)
    
    out_joint_df$UMAP1_cobolt <- out_cobolt_UMAP[rownames(out_joint_df),1]
    out_joint_df$UMAP2_cobolt <- out_cobolt_UMAP[rownames(out_joint_df),2]
    
    # perform multiVI and add their UMAP coordsin
    # note that similar to Cobolt MultiVI wants count data not logcounts
    out_multivi = multiVI_wrap(assay_list_counts[c("Multiome", "RNA", "ATAC")], verbose = TRUE)
    out_multivi_UMAP = calculateUMAP_rnames(out_multivi)
    
    out_joint_df$UMAP1_multivi <- out_multivi_UMAP[rownames(out_joint_df),1]
    out_joint_df$UMAP2_multivi <- out_multivi_UMAP[rownames(out_joint_df),2]
    
    
    
    lisi_values = compute_lisi(out_joint_corrected,
                               out_joint_df[rownames(out_joint_corrected),],
                               c("modality", "celltype"))
    lisi_values$probval <- probval
    lisi_values$numMultiome <- sum(modality == "Multiome")
    
    lisi_values_RNA = compute_lisi(out_joint_RNA_corrected,
                                   out_joint_df[rownames(out_joint_RNA_corrected),],
                                   c("modality", "celltype"))
    lisi_values$modality_RNA <- lisi_values_RNA$modality
    lisi_values$celltype_RNA <- lisi_values_RNA$celltype
    
    lisi_values_cobolt = compute_lisi(out_cobolt,
                                      out_joint_df[rownames(out_joint_corrected),],
                                      c("modality", "celltype"))
    lisi_values$modality_cobolt <- lisi_values_cobolt$modality
    lisi_values$celltype_cobolt <- lisi_values_cobolt$celltype
    
    lisi_values_multivi = compute_lisi(out_multivi,
                                       out_joint_df[rownames(out_joint_corrected),],
                                       c("modality", "celltype"))
    lisi_values$modality_multivi <- lisi_values_multivi$modality
    lisi_values$celltype_multivi <- lisi_values_multivi$celltype
    
    saveRDS(list(out_joint = out_joint,
                 out_joint_corrected = out_joint_corrected,
                 out_joint_RNA = out_joint_RNA,
                 out_joint_RNA_corrected = out_joint_RNA_corrected,
                 out_cobolt = out_cobolt,
                 out_multivi = out_multivi,
                 out_joint_df = out_joint_df,
                 lisi_values = lisi_values),
            file =  outFile)
  }
  
  lisi_values_list[[probval]] <- lisi_values
  
  # export some plots:
  g1 = ggplot(out_joint_df, aes(x = modality)) + 
    geom_bar(aes(fill = modality)) + 
    scale_fill_manual(values = modality_colours) +
    ylab("Number of cells") +
    theme_classic() +
    xlab("Simulated modality") +
    theme(legend.position = "none") +
    theme(axis.text = element_text(size = 15)) +
    theme(axis.title = element_text(size = 20)) +
    NULL
  g1
  ggsave(g1, file = paste0("../../Figures/raw/PBMC_nonoverlapping_multiome_", probval,
                           "_numCells.pdf"),
         width = 4, height = 4)
  
  g2 = mosaicDataUpSet(assay_list,
                       text.scale = c(3,1,2,1,3,3), 
                       point.size = 2.8, 
                       line.size = 1,
                       sets.x.label = "Cells",
                       mainbar.y.label = "Features",
                       main.bar.color = method_colours["StabMap"],
                       set_size.numbers_size = TRUE,
                       keep.order = TRUE,
                       sets = names(assay_list))
  g2
  pdf(paste0("../../Figures/raw/PBMC_nonoverlapping_multiome_", probval,
             "_upset.pdf"), 
      width = 5, height = 5)
  print(g2)
  dev.off()
  
  graph = mosaicDataTopology(assay_list)
  V(graph)$size = 50
  V(graph)$color = modality_colours[V(graph)$name]
  V(graph)$label.color = c("white", "black", "black")
  plot(graph, layout = layout.drl)
  
  pdf(paste0("../../Figures/raw/PBMC_nonoverlapping_multiome_", probval,
             "_network.pdf"), 
      width = 6, height = 6)
  plot(graph, layout = layout.drl)
  dev.off()
  
  g3 = ggplot(out_joint_df, aes(x = UMAP1_corrected, y = UMAP2_corrected)) + 
    geom_scattermore(aes(colour = modality), pointsize = 1.3, alpha = 0.5) +
    scale_colour_manual(values = modality_colours) +
    theme_classic() + 
    theme(legend.position = "bottom") +
    theme(axis.ticks = element_blank()) +
    theme(axis.text = element_blank()) +
    guides(colour = guide_legend(override.aes = list(size = 5, alpha = 1))) +
    coord_fixed() +
    xlab("UMAP1") + 
    ylab("UMAP2") +
    labs(colour = "") +
    NULL
  g3
  ggsave(g3 + theme(legend.position = "none"), file = paste0("../../Figures/raw/PBMC_nonoverlapping_multiome_", probval,
                                                             "_UMAP_modality.pdf"),
         height = 3, width = 3)
  
  g6 = as_ggplot(get_legend(g3))
  g6
  ggsave(g6, 
         file = paste0("../../Figures/raw/PBMC_nonoverlapping_multiome_", probval,
                       "_UMAP_modality_legend.pdf"),
         height = 1, width = 3)
  
  g3_RNA = ggplot(out_joint_df, aes(x = UMAP1_RNA, y = UMAP2_RNA)) + 
    geom_scattermore(aes(colour = modality), pointsize = 1.3, alpha = 0.5) +
    scale_colour_manual(values = modality_colours) +
    theme_classic() + 
    theme(legend.position = "bottom") +
    theme(axis.ticks = element_blank()) +
    theme(axis.text = element_blank()) +
    guides(colour = guide_legend(override.aes = list(size = 5, alpha = 1))) +
    coord_fixed() +
    xlab("UMAP1") + 
    ylab("UMAP2") +
    labs(colour = "") +
    NULL
  g3_RNA
  ggsave(g3_RNA + theme(legend.position = "none"), file = paste0("../../Figures/raw/PBMC_nonoverlapping_multiome_", probval,
                                                                 "_RNA_UMAP_modality.pdf"),
         height = 3, width = 3)
  
  g3_cobolt = ggplot(out_joint_df, aes(x = UMAP1_cobolt, y = UMAP2_cobolt)) + 
    geom_scattermore(aes(colour = modality), pointsize = 1.3, alpha = 0.5) +
    scale_colour_manual(values = modality_colours) +
    theme_classic() + 
    theme(legend.position = "bottom") +
    theme(axis.ticks = element_blank()) +
    theme(axis.text = element_blank()) +
    guides(colour = guide_legend(override.aes = list(size = 5, alpha = 1))) +
    coord_fixed() +
    xlab("UMAP1") + 
    ylab("UMAP2") +
    labs(colour = "") +
    NULL
  g3_cobolt
  ggsave(g3_cobolt + theme(legend.position = "none"), file = paste0("../../Figures/raw/PBMC_nonoverlapping_multiome_", probval,
                                                                    "_cobolt_UMAP_modality.pdf"),
         height = 3, width = 3)
  
  g3_multivi = ggplot(out_joint_df, aes(x = UMAP1_multivi, y = UMAP2_multivi)) + 
    geom_scattermore(aes(colour = modality), pointsize = 1.3, alpha = 0.5) +
    scale_colour_manual(values = modality_colours) +
    theme_classic() + 
    theme(legend.position = "bottom") +
    theme(axis.ticks = element_blank()) +
    theme(axis.text = element_blank()) +
    guides(colour = guide_legend(override.aes = list(size = 5, alpha = 1))) +
    coord_fixed() +
    xlab("UMAP1") + 
    ylab("UMAP2") +
    labs(colour = "") +
    NULL
  g3_multivi
  ggsave(g3_multivi + theme(legend.position = "none"), file = paste0("../../Figures/raw/PBMC_nonoverlapping_multiome_", probval,
                                                                     "_multivi_UMAP_modality.pdf"),
         height = 3, width = 3)
  
  library(RColorBrewer)
  colourCount = length(unique(out_joint_df$celltype))
  getPalette = colorRampPalette(brewer.pal(9,"Set1"))
  
  g4 = ggplot(out_joint_df, aes(x = UMAP1_corrected, y = UMAP2_corrected)) + 
    geom_scattermore(aes(colour = celltype), pointsize = 1.3, alpha = 0.5) +
    scale_colour_manual(values = getPalette(colourCount)) +
    theme_classic() + 
    theme(legend.position = "right",) +
    theme(axis.ticks = element_blank()) +
    theme(axis.text = element_blank()) +
    guides(colour = guide_legend(override.aes = list(size = 5, alpha = 1),
                                 ncol = 1)) +
    coord_fixed() +
    xlab("UMAP1") + 
    ylab("") +
    labs(colour = "") +
    NULL
  g4
  ggsave(g4 + theme(legend.position = "none"), 
         file = paste0("../../Figures/raw/PBMC_nonoverlapping_multiome_", probval,
                       "_UMAP_celltype.pdf"),
         height = 3, width = 3)
  
  g5 = as_ggplot(get_legend(g4))
  g5
  ggsave(g5, 
         file = paste0("../../Figures/raw/PBMC_nonoverlapping_multiome_", probval,
                       "_UMAP_celltype_legend.pdf"),
         height = 5, width = 4)
  
  
  g4_RNA = ggplot(out_joint_df, aes(x = UMAP1_RNA, y = UMAP2_RNA)) + 
    geom_scattermore(aes(colour = celltype), pointsize = 1.3, alpha = 0.5) +
    scale_colour_manual(values = getPalette(colourCount)) +
    theme_classic() + 
    theme(legend.position = "right",) +
    theme(axis.ticks = element_blank()) +
    theme(axis.text = element_blank()) +
    guides(colour = guide_legend(override.aes = list(size = 5, alpha = 1),
                                 ncol = 1)) +
    coord_fixed() +
    xlab("UMAP1") + 
    ylab("") +
    labs(colour = "") +
    NULL
  g4_RNA
  ggsave(g4_RNA + theme(legend.position = "none"), 
         file = paste0("../../Figures/raw/PBMC_nonoverlapping_multiome_", probval,
                       "_RNA_UMAP_celltype.pdf"),
         height = 3, width = 3)
  
  g4_cobolt = ggplot(out_joint_df, aes(x = UMAP1_cobolt, y = UMAP2_cobolt)) + 
    geom_scattermore(aes(colour = celltype), pointsize = 1.3, alpha = 0.5) +
    scale_colour_manual(values = getPalette(colourCount)) +
    theme_classic() + 
    theme(legend.position = "right",) +
    theme(axis.ticks = element_blank()) +
    theme(axis.text = element_blank()) +
    guides(colour = guide_legend(override.aes = list(size = 5, alpha = 1),
                                 ncol = 1)) +
    coord_fixed() +
    xlab("UMAP1") + 
    ylab("") +
    labs(colour = "") +
    NULL
  g4_cobolt
  ggsave(g4_cobolt + theme(legend.position = "none"), 
         file = paste0("../../Figures/raw/PBMC_nonoverlapping_multiome_", probval,
                       "_cobolt_UMAP_celltype.pdf"),
         height = 3, width = 3)
  
  g4_multivi = ggplot(out_joint_df, aes(x = UMAP1_multivi, y = UMAP2_multivi)) + 
    geom_scattermore(aes(colour = celltype), pointsize = 1.3, alpha = 0.5) +
    scale_colour_manual(values = getPalette(colourCount)) +
    theme_classic() + 
    theme(legend.position = "right",) +
    theme(axis.ticks = element_blank()) +
    theme(axis.text = element_blank()) +
    guides(colour = guide_legend(override.aes = list(size = 5, alpha = 1),
                                 ncol = 1)) +
    coord_fixed() +
    xlab("UMAP1") + 
    ylab("") +
    labs(colour = "") +
    NULL
  g4_multivi
  ggsave(g4_multivi + theme(legend.position = "none"), 
         file = paste0("../../Figures/raw/PBMC_nonoverlapping_multiome_", probval,
                       "_multivi_UMAP_celltype.pdf"),
         height = 3, width = 3)
  
  
  # Perform cell type classification across combinations and save the result:
  combs_raw = combn(c("RNA", "Multiome", "ATAC"), 2)
  combs = rbind(t(combs_raw), t(combs_raw[c(2,1),]))
  
  for (i in 1:nrow(combs)) {
    
    library(class)
    class_train = setNames(subset(out_joint_df, modality == combs[i,1])[,"celltype"], rownames(subset(out_joint_df, modality == combs[i,1])))
    data_train_multivi = out_multivi[names(class_train),]
    data_train_stabmap = out_joint_corrected[names(class_train),]
    data_train_cobolt = out_cobolt[names(class_train),]
    data_train_stabmap_RNA = out_joint_RNA_corrected[names(class_train),]
    
    class_test = setNames(subset(out_joint_df, modality == combs[i,2])[,"celltype"], rownames(subset(out_joint_df, modality == combs[i,2])))
    data_test_multivi = out_multivi[names(class_test),]
    data_test_stabmap = out_joint_corrected[names(class_test),]
    data_test_cobolt = out_cobolt[names(class_test),]
    data_test_stabmap_RNA = out_joint_RNA_corrected[names(class_test),]
    
    acc_multivi = mean(knn(data_train_multivi, data_test_multivi, class_train, k = 5) == class_test)
    acc_stabmap = mean(knn(data_train_stabmap, data_test_stabmap, class_train, k = 5) == class_test)
    acc_cobolt = mean(knn(data_train_cobolt, data_test_cobolt, class_train, k = 5) == class_test)
    acc_stabmap_RNA = mean(knn(data_train_stabmap_RNA, data_test_stabmap_RNA, class_train, k = 5) == class_test)
    
    pred_i = data.frame(
      probval = probval,
      prediction = paste0(combs[i,], collapse = "_"), 
      Method = c("MultiVI", "StabMap", "Cobolt", "StabMap_RNA"),
      acc = c(acc_multivi, acc_stabmap, acc_cobolt, acc_stabmap_RNA))
    
    pred = rbind(pred,pred_i)
  }
  
}
```

Make some plots - use local inverse simpson index to identify at which point there are too few multiome cells to adequately map data across.

```{r}
lisi_values_all = do.call(rbind, lisi_values_list)
lisi_values_all$probval <- factor(lisi_values_all$probval)
lisi_values_all$numMultiome <- factor(lisi_values_all$numMultiome)

saveRDS(lisi_values_all, file = "../../output/PBMC_Multiome_nonoverlapping_LISI_values.Rds")

vval = 3.5

g1 = ggplot(lisi_values_all, aes(x = numMultiome, y = modality, group = numMultiome)) +
  theme_classic() +
  xlab("") +
  theme(axis.text.x = element_blank()) +
  ylab("Modality LISI") +
  geom_boxplot(outlier.size = 0.1, fill = "gray75") + 
  geom_vline(xintercept = vval, linetype = "dotted") +
  theme(axis.title = element_text(size = 12)) +
  NULL
g1

g2 = ggplot(lisi_values_all, aes(x = numMultiome, y = celltype, group = numMultiome)) +
  theme_classic() +
  ylab("Cell type LISI") +
  xlab("Number of cells in Multiome dataset") +
  geom_boxplot(outlier.size = 0.1, fill = "gray75") + 
  geom_vline(xintercept = vval, linetype = "dotted") +
  theme(axis.title = element_text(size = 12)) +
  NULL

g = g1 + g2 + plot_layout(nrow = 2)
g

ggsave(g, file = "../../Figures/raw/PBMC_nonoverlapping_multiome_LISI.pdf",
       height = 3, width = 8)
```

# Predict cell type labels across datasets

Use one set of cells to predict another set of cells cell types and calculate accuracy using true
labels

```{r}
num_map = lisi_values_all[,c("probval","numMultiome")]
num_map <- num_map[!duplicated(num_map),]
num_map_val <- setNames(num_map$numMultiome, num_map$probval)

pred$numMultiome <- num_map_val[as.character(pred$probval)]

g = ggplot(pred, aes(x = numMultiome, y = acc)) + 
  geom_point(aes(colour = Method), alpha = 0.5) + 
  geom_smooth(aes(group = Method, colour = Method), fill = NA) +
  theme_classic() +
  theme(legend.position = "none") +
  xlab("Number of cells in Multiome dataset") +
  ylab("Cell type accuracy")
g

ggsave(g, file = "../../Figures/raw/PBMC_nonoverlapping_accuracy.pdf",
       height = 4, width = 5)

g_leg = as_ggplot(get_legend(g + theme(legend.position = "bottom") + guides(colour = guide_legend(title = ""))))
g_leg
ggsave(g_leg, 
       file = paste0("../../Figures/raw/PBMC_nonoverlapping_accuracy_legend.pdf"),
       height = 1, width = 5)
```

Finish

```{r}
sessionInfo()
```
