---
title: "StabMap: Generate PBMC CYTOF data object"
author: "Shila Ghazanfar"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
       html_document:
                     toc: true
                     toc_float:
                           collapsed: false
                           smooth_scroll: false
                     code_folding: hide
                     fig_width: 10 
                     fig_height: 8
---

<!-- TODO- Use ABnames package from Helen Lindsay (Raphael Gottardo group) to  -->
<!-- match the antibody names -->
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE,
                      cache = FALSE, cache.lazy = FALSE)
```

```{r}
set.seed(2021)
```

Load scripts and packages.

```{r}
library(HDCytoData)
source("../scripts/initialise.R")
```

Load data from Bodenmiller (2012) via the HDCytoData package. This dataset 
includes two conditions of stimulated and unstimulated PBMCs from healthy 
individuals.

```{r}
d_SE = Bodenmiller_BCR_XL_SE(metadata = FALSE)
d_SE
```

Perform some exploration of this data.

```{r}
dim(d_SE)
assay(d_SE[1:5,1:5], "exprs")
colnames(d_SE)
head(rowData(d_SE))
table(rowData(d_SE)$group_id)
table(rowData(d_SE)$group_id, rowData(d_SE)$patient_id)
```

Extract the CYTOF data matrix, transpose to match typical use in scRNA-seq
studies. Remove the variables Time, Cell_length, DNA-1 and DNA-2.

```{r}
vars_to_keep = rownames(subset(colData(d_SE), marker_class != "none"))
exprs_cytof = t(assay(d_SE, "exprs")[,vars_to_keep])
class(exprs_cytof)
dim(exprs_cytof)

sce.cytof = SingleCellExperiment(assays = list(counts = exprs_cytof))
```

Add cell metadata

```{r}
meta = rowData(d_SE)

colData(sce.cytof) <- meta
```

Perform some exploration of this data, first perform transformation as 
suggested by the vignette, i.e. asinh with cofactor 5.

```{r}
# Normalisation
logcounts(sce.cytof) <- asinh(counts(sce.cytof)/5)

# PCA
sce.cytof <- scater::runPCA(sce.cytof, ncomponents = 10)

# UMAP
sce.cytof <- scater::runUMAP(sce.cytof, dimred="PCA", n_neighbors = 25, min_dist = 0.3)

plotPCA(sce.cytof, point_size=0.5, point_alpha=1, colour_by = "group_id")

plotUMAP(sce.cytof, point_size=0.5, point_alpha=1)
plotUMAP(sce.cytof, point_size=0.5, point_alpha=1, colour_by = "population_id")
plotUMAP(sce.cytof, point_size=0.5, point_alpha=1, colour_by = "sample_id")
plotUMAP(sce.cytof, point_size=0.5, point_alpha=1, colour_by = "patient_id")
plotUMAP(sce.cytof, point_size=0.5, point_alpha=1, colour_by = "group_id")
```

Save the object as RDS.

```{r}
# save the object as RDS
saveRDS(sce.cytof, file = "../../CYTOF/cytof_PBMC_sce.Rds")
```

# Finish

```{r}
sessionInfo()
```
